---
title: "Planning optimal fire mosaics- initial findings"
author: "Chris Adlam"
date: "10/2018"
output:
  pdf_document: default
  html_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, error = F) 
knitr::knit_hooks$set(
   error = function(x, options) {
     paste('\n\n<div class="alert alert-danger">',
           gsub('##', '\n', gsub('^##\ Error', '**Error**', x)),
           '</div>', sep = '\n')
   },
   warning = function(x, options) {
     paste('\n\n<div class="alert alert-warning">',
           gsub('##', '\n', gsub('^##\ Warning:', '**Warning**', x)),
           '</div>', sep = '\n')
   },
   message = function(x, options) {
     paste('\n\n<div class="alert alert-info">',
           gsub('##', '\n', x),
           '</div>', sep = '\n')
   }
)
```

*************************

# Overview
**Project aims:**   

*   Evaluate the effect of fire severity and time (and, conversely, lack of fire) on native plant, bird, lichen and insect communities in the WKRP area;    

*   Suggest proportions of each habitat on the landscape that would maximize species diversity;    
*   Make management recommendations for forest restoration treatments that will be most likely to maximize biodiversity.  
  
**Methods:**    
Surveys were conducted at 48 sites divided into long-unburnt (hereafter UN), low severity (hereafter LS) and high severity (hereafter HS) fire, ranging from 4 to 31 years since the last fire. No repeat burns were studied.  
  
Sampling methods involved: surveys of all native plants and lichens (growing on trees) within an 11.3m radius circle; two 10-minute bird point counts; one flight-intercept trap left for 2 days to passively collect flying insects. Preliminary data for plants and birds are presented here. Lichen and insect specimens are still being identified (completion expected by December 2018).

```{r include = F}
#library(lme4)
#library(lmerTest)
#library(pbkrtest)
#library(car)
#library(reshape2)
#library(splitstackshape)
#library(magrittr)
#library(readr)
#library(viridis)
#library(purrr)
#library(tibble)
#library(broom)
#library(indicspecies)
#library(permute)
#library(perm)
#library(magrittr)
#library(knitr)
#knit2html("Fire_mosaics.html")
source("scripts/splist2presabs.R")
source("scripts/data_load.R")
source("scripts/pairwise_permanova.R")


```


```{r include=F}
#obsolete?
# To get p/a:
plant_dat_pa_l <- splist2presabs(plant_dat, sites.col = 1, sp.col = 2) %>% 
  gather(key = species, value = pa, ABCO:YAMI) #ABCO:YAMI for all native herbs in all sites

# The following is to remove species detected too few times
# convert to wide format for following analysis
plant_matrix1 <- spread(data = plant_dat, key = species, value = cover, fill = 0)

# optional/incomplete:remove plants with few sightings; not sure this makes much difference
## Here I'm making a row for the frequency of detection
plant_dat$cover[plant_dat$cover > 0] <- 1 
plant_matrix2 <- spread(data = plant_dat, key = species, value = cover, fill = 0) %>% 
  bind_rows(summarise_all(., funs(if(is.numeric(.)) sum(.) else "Total")))

# adding dummy species (eg. Webster 2010)
plant_matrix <- plant_matrix1 %>%
  bind_rows(., plant_matrix2[111,]) %>% #for only 2018 data, change 111 to 49
  mutate("DUMB" = "0.5") 

## keep only columns where minimum is reached (if subsetting some sites, change row number, here 49) (true)
plant_matrix <- plant_matrix[, (plant_matrix[49, ]) > 5]
## adding back in the site id and removing true/false row (and making sure it's all read in as numeric)
plant_matrix <- plant_matrix[-111, ] %>% #for only 2018 data, use change 111 to 49
  mutate_if(is.character, as.numeric)

# Make Bray-Curtis (?) dissimilarity matrix
#plants_matrix <- as.matrix(plant_matrix1[, -1])

#this used to be plant_dat_cov, same as plant_mrpp_d
plant_dat_cov_w <- left_join(plant_matrix1, site_data, by = "site_id") %>% 
  mutate(sev_tsf = paste(sev, tsf_cat, sep = "-"))

plant_dat_cov_w$tsf_cat <- as.factor(plant_dat_cov_w$tsf_cat)
```

# Results
```{r}
# This chunk is about comparing the frequency of different species in each habitat (see Wayman 2007)

# Calculating the number of occurrences in each trt
# First get the data matrices for each habitat type
plant_mat_UNLS <- plant_mat_pa_w %>%
  left_join(site_data, by = "site_id") %>% 
  filter(sev == "u" | sev == "l") %>% 
  dplyr::select(ABCO:YAMI)

plant_mat_UN <- plant_mat_pa_w %>%
  left_join(site_data, by = "site_id") %>% 
  filter(sev == "u") %>% 
  dplyr::select(ABCO:YAMI)

plant_mat_LS <- plant_mat_pa_w %>%
  left_join(site_data, by = "site_id") %>% 
  filter(sev == "l") %>% 
  dplyr::select(ABCO:YAMI)

plant_mat_HS <- plant_mat_pa_w %>% 
  left_join(site_data, by = "site_id") %>% 
  filter(sev=="h") %>% 
  dplyr::select(ACMA:YAMI)

plant_mat_MULT <- plant_mat_pa_w %>%
  left_join(site_data, by = "site_id") %>% 
  filter(sev == "multiple") %>% 
  dplyr::select(ABCO:YAMI)

plant_mat_Rx <- plant_mat_pa_w %>%
  left_join(site_data, by = "site_id") %>% 
  filter(sev == "Rx") %>% 
  dplyr::select(ABCO:YAMI)

#Getting the number of detections for each species
plant_dat_sum_HS <- rbind(plant_mat_HS[,-1], colSums(plant_mat_HS[-1,-1]))
plant_sum_HS <- data.frame(t(tail(plant_dat_sum_HS, n = 1))) # keep only last row (sum) and transpose
plant_sum_HS <- cbind(rownames(plant_sum_HS), data.frame(plant_sum_HS, row.names=NULL)) # row names to column
colnames(plant_sum_HS) <- c("species", "sum")

plant_dat_sum_UNLS <- rbind(plant_mat_UNLS[,-1], colSums(plant_mat_UNLS[-1,-1]))
plant_sum_UNLS <- data.frame(t(tail(plant_dat_sum_UNLS, n= 1))) 
plant_sum_UNLS <- cbind(rownames(plant_sum_UNLS), data.frame(plant_sum_UNLS, row.names=NULL))
colnames(plant_sum_UNLS) <- c("species", "sum")

plant_dat_sum_LS <- rbind(plant_mat_LS[,-1], colSums(plant_mat_LS[-1,-1]))
plant_sum_LS <- data.frame(t(tail(plant_dat_sum_LS, n=1))) 
plant_sum_LS <- cbind(rownames(plant_sum_LS), data.frame(plant_sum_LS, row.names=NULL))
colnames(plant_sum_LS) <- c("species", "sum")

plant_dat_sum_UN <- rbind(plant_mat_UN[,-1], colSums(plant_mat_UN[-1,-1]))
plant_sum_UN <- data.frame(t(tail(plant_dat_sum_UN, n = 1))) 
plant_sum_UN <- cbind(rownames(plant_sum_UN), data.frame(plant_sum_UN, row.names=NULL))
colnames(plant_sum_UN) <- c("species", "sum")

plant_dat_sum_MULT <- rbind(plant_mat_MULT[,-1], colSums(plant_mat_MULT[-1,-1]))
plant_sum_MULT <- data.frame(t(tail(plant_dat_sum_MULT, n = 1))) 
plant_sum_MULT <- cbind(rownames(plant_sum_MULT), data.frame(plant_sum_MULT, row.names=NULL))
colnames(plant_sum_MULT) <- c("species", "sum")

plant_dat_sum_Rx <- rbind(plant_mat_Rx[,-1], colSums(plant_mat_Rx[-1,-1]))
plant_sum_Rx <- data.frame(t(tail(plant_dat_sum_Rx, n = 1))) 
plant_sum_Rx <- cbind(rownames(plant_sum_Rx), data.frame(plant_sum_Rx, row.names=NULL))
colnames(plant_sum_Rx) <- c("species", "sum")

plant_dat_sum <- plant_sum_UNLS %>% 
  left_join(plant_sum_HS, by = "species") %>% 
  left_join(plant_sum_UN, by = "species") %>% 
  left_join(plant_sum_LS, by = "species") %>% 
  left_join(plant_sum_MULT, by = "species") %>% 
  left_join(plant_sum_Rx, by = "species")

colnames(plant_dat_sum) <- c("species", "det_UNLS", "det_HS", "det_UN", "det_LS", "det_MULT", "det_Rx")
plant_dat_sum[is.na(plant_dat_sum)] <- 0

# Now dividing number of detections by number of sites and turning into %
plant_dat_percent <- plant_dat_sum %>% 
  mutate(., UNLS_occ = (det_UNLS/(nrow(plant_dat_sum_UNLS) -1)*100)) %>% 
  mutate(., HS_occ = (det_HS/(nrow(plant_dat_sum_HS) -1)*100)) %>% 
  mutate(., UN_occ = (det_UN/(nrow(plant_dat_sum_UN) -1)*100)) %>% 
  mutate(., LS_occ = (det_LS/(nrow(plant_dat_sum_LS) -1)*100)) %>% 
  mutate(., MULT_occ = (det_MULT/(nrow(plant_dat_sum_MULT) -1)*100)) %>% 
  mutate(., Rx_occ = (det_Rx/(nrow(plant_dat_sum_Rx) -1)*100)) %>% 
  mutate(., total_occ = (det_UNLS + det_HS + det_MULT + det_Rx)) %>% 
  filter(total_occ >5)  # keep only species detected more than 5 times

print_occ_table <- plant_dat_percent[, -c(2:7)]

write.csv(print_occ_table, file = "output/print_occ_table.csv")

# do this again with %cov!

# First get the data matrices for each cover cat
plant_mat_1 <- plant_dat_pa_w %>%
#  left_join(subset(plant_dat_pa_w, select = site_id:YAMI), by = "site_id") %>% 
  filter(cov_cat == "1") %>% 
  dplyr::select(ACAM:YAMI)

plant_mat_2 <- plant_dat_pa_w %>%
#  left_join(subset(plant_dat_pa_w, select = site_id:YAMI), by = "site_id") %>% 
  filter(cov_cat == "2") %>% 
  dplyr::select(ACAM:YAMI)

plant_mat_3 <- plant_dat_pa_w %>%
#  left_join(plant_dat_pa_w[,-c(2:213)], by = "site_id") %>% 
  filter(cov_cat == "3") %>% 
  dplyr::select(ACAM:YAMI)

#Getting the number of detections for each species
plant_dat_sum_1 <- rbind(plant_mat_1[,-1], colSums(plant_mat_1[-1,-1]))
plant_sum_1 <- data.frame(t(tail(plant_dat_sum_1, n = 1))) # keep only last row (sum) and transpose
plant_sum_1 <- cbind(rownames(plant_sum_1), data.frame(plant_sum_1, row.names=NULL)) # row names to column
colnames(plant_sum_1) <- c("species", "sum")

plant_dat_sum_2 <- rbind(plant_mat_2[,-1], colSums(plant_mat_2[-1,-1]))
plant_sum_2 <- data.frame(t(tail(plant_dat_sum_2, n= 1))) 
plant_sum_2 <- cbind(rownames(plant_sum_2), data.frame(plant_sum_2, row.names=NULL))
colnames(plant_sum_2) <- c("species", "sum")

plant_dat_sum_3 <- rbind(plant_mat_3[,-1], colSums(plant_mat_3[-1,-1]))
plant_sum_3 <- data.frame(t(tail(plant_dat_sum_3, n=1))) 
plant_sum_3 <- cbind(rownames(plant_sum_3), data.frame(plant_sum_3, row.names=NULL))
colnames(plant_sum_3) <- c("species", "sum")

plant_dat_sum <- plant_sum_1 %>% 
  left_join(plant_sum_2, by = "species") %>% 
  left_join(plant_sum_3, by = "species")

colnames(plant_dat_sum) <- c("species", "det_1", "det_2", "det_3")
plant_dat_sum[is.na(plant_dat_sum)] <- 0

# Now dividing number of detections by number of sites and turning into %
plant_dat_percent <- plant_dat_sum %>% 
  mutate(., occ1 = (det_1/(nrow(plant_dat_sum_1) -1)*100)) %>% 
  mutate(., occ2 = (det_2/(nrow(plant_dat_sum_2) -1)*100)) %>% 
  mutate(., occ3 = (det_3/(nrow(plant_dat_sum_3) -1)*100)) %>% 
  mutate(., total_occ = (det_1 + det_2 + det_3)) %>% 
  filter(total_occ >5)  # keep only species detected more than 5 times

print_occ_table_cov <- plant_dat_percent[, -c(2:4)]

write.csv(print_occ_table_cov, file = "output/print_occ_table_cov.csv")

# species that are found at least 2x more in HS
plant_dat_cov1 <- plant_dat_sum %>% 
  mutate(ratio = cov_HS/cov_UNLS)# %>% 
#  filter(ratio >= 2)

# species that are found at least 2x more in HS
plant_dat_cov2 <- plant_dat_sum %>% 
  mutate(ratio = cov_HS/cov_UNLS)# %>% 
#  filter(ratio <= 0.5)

cov_tbl <- plant_mat_cov_w %>% 
  rbind(plant_dat_cov2)

pa_tbl <- plant_dat_sum1 %>% 
  rbind(plant_dat_sum2)

pa_cov_tbl <- plant_dat_cov1 %>% 
  full_join(plant_dat_sum1, by = "species") %>% 
  filter(!is.na(sum_HS)) %>% 
  filter_at(vars(ratio.x, ratio.y), all_vars(.>1.5 | .<0.5))

write.csv(pa_cov_tbl, file = "output/pa_cov_tbl_complete.csv")


### cov (the end part)

plant_dat_sum <- plant_sum_UNLS %>% 
  left_join(plant_sum_HS, by = "species")
colnames(plant_dat_sum) <- c("species", "cov_UNLS", "cov_HS")

# species that are found at least 2x more in HS
plant_dat_cov1 <- plant_dat_sum %>% 
  mutate(ratio = cov_HS/cov_UNLS)# %>% 
#  filter(ratio >= 2)

# species that are found at least 2x more in HS
plant_dat_cov2 <- plant_dat_sum %>% 
  mutate(ratio = cov_HS/cov_UNLS)# %>% 
#  filter(ratio <= 0.5)

cov_tbl <- plant_mat_cov_w %>% 
  rbind(plant_dat_cov2)

pa_tbl <- plant_dat_sum1 %>% 
  rbind(plant_dat_sum2)

pa_cov_tbl <- plant_dat_cov1 %>% 
  full_join(plant_dat_sum1, by = "species") %>% 
  filter(!is.na(sum_HS)) %>% 
  filter_at(vars(ratio.x, ratio.y), all_vars(.>1.5 | .<0.5))

write.csv(pa_cov_tbl, file = "output/pa_cov_tbl_complete.csv")

```

```{r}
# Birds
# tree cover categories

# First get the data matrices for each cover cat
bird_mat_1 <- bird_dat_w %>%
#  left_join(subset(bird_dat_pa_w, select = site_id:YAMI), by = "site_id") %>% 
  filter(cov_cat == "1") %>% 
  dplyr::select(ACWO:YRWA)

bird_mat_2 <- bird_dat_w %>%
#  left_join(subset(bird_dat_pa_w, select = site_id:YAMI), by = "site_id") %>% 
  filter(cov_cat == "2") %>% 
  dplyr::select(ACWO:YRWA)

bird_mat_3 <- bird_dat_w %>%
#  left_join(bird_dat_pa_w[,-c(2:213)], by = "site_id") %>% 
  filter(cov_cat == "3") %>% 
  dplyr::select(ACWO:YRWA)

#Getting the number of detections for each species
bird_dat_sum_1 <- rbind(bird_mat_1[,-1], colSums(bird_mat_1[-1,-1]))
bird_sum_1 <- data.frame(t(tail(bird_dat_sum_1, n = 1))) # keep only last row (sum) and transpose
bird_sum_1 <- cbind(rownames(bird_sum_1), data.frame(bird_sum_1, row.names=NULL)) # row names to column
colnames(bird_sum_1) <- c("species", "sum")

bird_dat_sum_2 <- rbind(bird_mat_2[,-1], colSums(bird_mat_2[-1,-1]))
bird_sum_2 <- data.frame(t(tail(bird_dat_sum_2, n= 1))) 
bird_sum_2 <- cbind(rownames(bird_sum_2), data.frame(bird_sum_2, row.names=NULL))
colnames(bird_sum_2) <- c("species", "sum")

bird_dat_sum_3 <- rbind(bird_mat_3[,-1], colSums(bird_mat_3[-1,-1]))
bird_sum_3 <- data.frame(t(tail(bird_dat_sum_3, n=1))) 
bird_sum_3 <- cbind(rownames(bird_sum_3), data.frame(bird_sum_3, row.names=NULL))
colnames(bird_sum_3) <- c("species", "sum")

bird_dat_sum <- bird_sum_1 %>% 
  left_join(bird_sum_2, by = "species") %>% 
  left_join(bird_sum_3, by = "species")

colnames(bird_dat_sum) <- c("species", "det_1", "det_2", "det_3")
bird_dat_sum[is.na(bird_dat_sum)] <- 0

# Now dividing number of detections by number of sites and turning into %
bird_dat_percent <- bird_dat_sum %>% 
  mutate(., occ1 = (det_1/(nrow(bird_dat_sum_1) -1)*100)) %>% 
  mutate(., occ2 = (det_2/(nrow(bird_dat_sum_2) -1)*100)) %>% 
  mutate(., occ3 = (det_3/(nrow(bird_dat_sum_3) -1)*100)) %>% 
  mutate(., total_occ = (det_1 + det_2 + det_3)) %>% 
  filter(total_occ >5)  # keep only species detected more than 5 times

print_occ_table_cov <- bird_dat_percent[, -c(2:4)]

write.csv(print_occ_table_cov, file = "output/print_occ_table_BIRDS_cov.csv")



#site_data %>% group_by(sev) %>% summarise(mean = mean(tree_cov), SD = sd(tree_cov), freq = n())


# species that are found at least 2x more in HS
bird_dat_cov1 <- bird_dat_sum %>% 
  mutate(ratio = cov_HS/cov_UNLS)# %>% 
#  filter(ratio >= 2)

# species that are found at least 2x more in HS
bird_dat_cov2 <- bird_dat_sum %>% 
  mutate(ratio = cov_HS/cov_UNLS)# %>% 
#  filter(ratio <= 0.5)

cov_tbl <- bird_mat_cov_w %>% 
  rbind(bird_dat_cov2)

pa_tbl <- bird_dat_sum1 %>% 
  rbind(bird_dat_sum2)

pa_cov_tbl <- bird_dat_cov1 %>% 
  full_join(bird_dat_sum1, by = "species") %>% 
  filter(!is.na(sum_HS)) %>% 
  filter_at(vars(ratio.x, ratio.y), all_vars(.>1.5 | .<0.5))

write.csv(pa_cov_tbl, file = "output/BIRDS_cov_tbl_complete.csv")
```


```{r}
#Obsolete?
# Calculating the frequency in each trt
plant_glm_pa1 <- plant_glm_pa %>% 
  dplyr::select(site_id:WOFI)

plant_glm_pa1$site_id <- as.numeric(as.character(plant_glm_pa1$site_id))

plant_glm_pa1 <- rbind(plant_glm_pa1, colSums(plant_glm_pa1[-1,])) %>%
  select_if(function(.) last(.) > 4) # keep only species detected 5 or more times

plant_glm_pa1$site_id <- as.factor(plant_glm_pa1$site_id)

plant_mat_UNLS <- plant_glm_pa1 %>%
  left_join(site_data, by = "site_id") %>% 
  filter(sev!="h") %>% 
  dplyr::select(ACMA:WHMO)

plant_mat_HS <- plant_glm_pa1 %>% 
  left_join(site_data, by = "site_id") %>% 
  filter(sev=="h") %>% 
  dplyr::select(ACMA:WHMO)

plant_dat_sum_HS <- rbind(plant_mat_HS[,-1], colSums(plant_mat_HS[-1,-1]))
plant_sum_HS <- data.frame(t(plant_dat_sum_HS[-c(1:23),])) 
plant_sum_HS <- cbind(rownames(plant_sum_HS), data.frame(plant_sum_HS, row.names=NULL))
colnames(plant_sum_HS) <- c("species", "sum")

plant_dat_sum_UNLS <- rbind(plant_mat_UNLS[,-1], colSums(plant_mat_UNLS[-1,-1]))
plant_sum_UNLS <- data.frame(t(plant_dat_sum_UNLS[-c(1:25),])) 
plant_sum_UNLS <- cbind(rownames(plant_sum_UNLS), data.frame(plant_sum_UNLS, row.names=NULL))
colnames(plant_sum_UNLS) <- c("species", "sum")

plant_dat_sum <- plant_sum_UNLS %>% 
  left_join(plant_sum_HS, by = "species")
colnames(plant_dat_sum) <- c("species", "sum_UNLS", "sum_HS")

# species that are found at least 2x more in HS
plant_dat_sum1 <- plant_dat_sum %>% 
  mutate(ratio = sum_HS/sum_UNLS)# %>% 
#  filter(ratio >= 2)

# species that are found at least 2x more in HS
plant_dat_sum2 <- plant_dat_sum %>% 
  mutate(ratio = sum_HS/sum_UNLS)# %>% 
#  filter(ratio <= 0.5)

```

```{r}
#Calculatign average and SD of tree cover for each sp

species_cov1 <- plant_dat_pa_w %>% 
  mutate_at(vars(ABCO:YAMI),
            function(x) ifelse(x == "1", .$tree_cov, x))

species_cov2 <- as.data.frame(t(species_cov1[,c(2:182)]))
species_cov2[species_cov2 == 0] <- NA

#install.packages("matrixStats")
#library(matrixStats)
#RowSD <- function(x) {
#  sqrt(rowSums((x - rowMeans(x))^2)/(dim(x)[2] - 1))
#}

species_cov <- species_cov2 %>% 
#  rowwise() %>%
  mutate(mean_cov = rowMeans(.[, c(1:110)], na.rm=TRUE), sd_cov = apply(.[(1:110)], na.rm=TRUE ,1,sd))

species_cov <- cbind(species_cov, species = rownames(species_cov2)) %>% 
  select(., species, mean_cov, sd_cov) %>% 
  filter(., !is.na(sd_cov))


#
plant_dat_cov_l_tr_cov <- plant_dat_cov_l %>% 
  left_join(site_data, by = "site_id") %>% 
  filter(cover != 0) %>% 
  select(site_id, species, tree_cov, sev)

unique(plant_dat_cov_l_tr_cov$sev)

library(cowplot)
p <- ggplot(plant_dat_cov_l_tr_cov, aes(x = reorder(species, tree_cov, FUN = median), y = tree_cov)) +
  geom_boxplot() +
  geom_jitter(aes(color = sev)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10))

p

#write.csv(species_cov, file = "output/mean_sd_COV_perSP.csv")
  
```


### NMDS

####plants

Non-metric multidimensional scaling (NMDS) is a way to graphically represent the difference between sampling sites. In the following graph, each point represents a site, and the distance between two points is proportional to the difference in the species composition at each site. The axes don't necessarily reflect ecologically meaningful factors.  

```{r, echo=FALSE,results='hide',fig.keep='all'}
#test NMDS method https://stackoverflow.com/questions/14711470/plotting-envfit-vectors-vegan-package-in-ggplot2

#library(grid)

#sub <- c("6", "25","27","28","51","32","47","49","37","41","40","52","53","1","30","33","34","38")

#plant_matrixx <- plant_matrix %>%
#  filter(site_id %in% sub) %>% 
#  mutate(group = ifelse(site_id %in% c("6", "25","27","28","51","32","47","49","37","41","40"), #1, 2))

#nmds_dat <- left_join(plant_mat_pa_w, lichen_mat_genus_pa_w, by = "site_id")


plant_dat_filt <- plant_dat_cov_w %>% filter((tsf < 25 & tsf > 4) | sev=="u")
nmds_dat <- plant_dat_filt %>% select(ABCO:YAMI)
nmds_dat[is.na(nmds_dat)] <- 0

# calculate distance for NMDS
NMDS.log<- log1p(nmds_dat)
                
#                [,-c(1,20)]) #previously tried log1p transformation for 2018, but with both years the stress is too high

#NMDS
sol <- metaMDS(NMDS.log, k = 3)

# scrs is the base for the plot
scrs <- as.data.frame(scores(sol, display = "sites"))
scrs <- cbind(scrs, site_id = plant_dat_filt$site_id)
scrs <- merge(scrs, plant_dat_filt, by = "site_id")

# To only show statistically significant species with arrows:
ef<-envfit(sol, NMDS.log, perm=999)
ef.df<-as.data.frame(ef$vectors$arrows*sqrt(ef$vectors$r))
ef.df$species<-rownames(ef.df)

#only significant pvalues
#shortcutting ef$vectors
A <- as.list(ef$vectors)
#creating the dataframe
pvals<-as.data.frame(A$pvals)
arrows<-as.data.frame(A$arrows*sqrt(A$r))
C<-cbind(arrows, pvals)
#subset
Cred<-subset(C,pvals<0.05)
Cred <- cbind(Cred, species = rownames(Cred))# %>% 
#  merge(., plant_names, by="species")

#aspect contour plot (testing)
#plot(sol,  display = "sites", main = "aspect contour")
#ordisurf(NMDS_test,Elevation, main="Elevation contour" ,col="forestgreen")
#orditorp(NMDS_test,display="species",col="grey30",air=0.1,cex=1)
#text(NMDS_test,display="species",col="grey30", cex=1)

#scrs_f <- scrs %>% 
#  dplyr::select(NMDS1, NMDS2, aspect)

#o <- ordisurf(sol, scrs$aspect)

# this bit is redundant (Cred = spp.scrs)
#vf <- envfit(sol, NMDS.log, perm = 999)
#spp.scrs <- as.data.frame(scores(vf, display = "vectors"))
#spp.scrs <- cbind(spp.scrs, Species = rownames(spp.scrs))
#spp.scrs <- cbind(spp.scrs, pvals)
#spp.scrs <- subset(spp.scrs,pvals<0.05)

# now envfit for env factors
scrs_filt <- scrs %>% 
  dplyr::select(site_id, folded_aspect, elev, heat_load, tree_cov, shrub_cov, bare_cov, herb_cov, cov_cat)

sol2 <-metaMDS((scrs_filt[,-(1:4)]))

fit <- envfit(sol2, scrs_filt[,-(1:4)], na.rm = T, perm = 999)
ef.df1<-as.data.frame(fit$vectors$arrows*sqrt(fit$vectors$r))
ef.df1$species<-rownames(ef.df1)


#only significant pvalues
#shortcutting ef$vectors
A1 <- as.list(fit$vectors)
#creating the dataframe
pvals1<-as.data.frame(A1$pvals)
arrows1<-as.data.frame(A1$arrows*sqrt(A1$r))
C1<-cbind(arrows1, pvals1)
C1 <- cbind(C1, Env_factors = rownames(C1))
#subset
Cred1<-subset(C1,pvals1<0.05)
Cred1 <- cbind(Cred1, Env_factors = rownames(Cred1))

#Axis 1 and 2
p1 <- ggplot(scrs, aes(x = NMDS1, y = NMDS2, colour = sev)) +
  geom_point() +
  scale_color_manual(values=c("#FF0000", "#3399FF", "#FFFF00", "#FF9933", "#993300", "#33FF00")) +
  coord_fixed() + ## need aspect ratio of 1!  
  stat_ellipse() +
#  geom_segment(data = Cred, #arrows for species
#               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
#               arrow = arrow(length = unit(0.25, "cm")), colour = "grey") +
#  geom_segment(data = C1,
#               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
#               arrow = arrow(length = unit(0.25, "cm")), colour = "red") +
#  geom_text(data = Cred, aes(x = NMDS1, y = NMDS2, label = species),
#            size = 3) + #species names at end of arrows
#  geom_text_repel(data = C1, aes(x = NMDS1, y = NMDS2, label = Env_factors),
#            size = 3, colour = "red") +
  labs(title = "Figure 1: NMDS for tree cov")

p1

#Axis 2 and 3
p2 <- ggplot(scrs) +
  geom_point(mapping = aes(x = NMDS2, y = NMDS3, colour = sev)) +
    scale_color_manual(values=c("#FF0000", "#3399FF", "#FFFF00", "#FF9933", "#993300", "#33FF00")) +
  coord_fixed() + ## need aspect ratio of 1!  
#  geom_segment(data = Cred,
#               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
#               arrow = arrow(length = unit(0.25, "cm")), colour = "grey") +
#  geom_segment(data = C1,
#               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
#               arrow = arrow(length = unit(0.25, "cm")), colour = "red") +
#  geom_text(data = Cred, aes(x = NMDS1, y = NMDS2, label = species),
#            size = 3) +
#  geom_text_repel(data = C1, aes(x = NMDS1, y = NMDS2, label = Env_factors),
#            size = 3, colour = "red") +
  labs(title = "Figure 2: NMDS for all native plants (2&3)")

p2


#Axis 1 and 3
p3 <- ggplot(scrs) +
  geom_point(mapping = aes(x = NMDS1, y = NMDS3, colour = sev)) +
    scale_color_manual(values=c("#FF0000", "#3399FF", "#FFFF00", "#FF9933", "#993300", "#33FF00")) +
  coord_fixed() + ## need aspect ratio of 1!  
#  geom_segment(data = Cred,
#               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
#               arrow = arrow(length = unit(0.25, "cm")), colour = "grey") +
#  geom_segment(data = C1,
#               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
#               arrow = arrow(length = unit(0.25, "cm")), colour = "red") +
#  geom_text(data = Cred, aes(x = NMDS1, y = NMDS2, label = species),
#            size = 3) +
#  geom_text_repel(data = C1, aes(x = NMDS1, y = NMDS2, label = Env_factors),
#            size = 3, colour = "red") +
  labs(title = "Figure 3: NMDS for all native plants (1&3)")

p3

p <- plot_grid(p1, p2, p3)

p

p3d <- plot_ly(scrs, x = ~NMDS1, y = ~NMDS2, z = ~NMDS3, color = ~cov_cat, symbol = ~sev, symbols = (c("square", "diamond", "circle", "cross", "x")), text = ~paste('cc: ', tree_cov)) %>%
  add_markers() %>%
  layout(scene = list(xaxis = list(title = 'NMDS1'),
                     yaxis = list(title = 'NMDS2'),
                     zaxis = list(title = 'NMDS3')))
p3d

#colors = c("#FF0000", "#3399FF", "#FFFF00", "#FF9933")
# , "#993300", "#33FF00"
#text = ~paste('site: ', scrs$site_id)


#  stat_contour(data = ordi.plant.na, aes(x, y, z = z)) + #can change the binwidth depending on how many contours you
#  scale_colour_manual("Post-fire habitats",  values = c("red", "green", "blue")) + #sets the name of the legend for shape, and says which symbols we want (equivalent to the pch command in plot)
#  labs(colour = "Severity")+ #another way to set the labels, in this case, for the colour legend
#  scale_colour_gradient(high = "darkgreen", low = "darkolivegreen1") + #here we set the high and low of the colour scale.  Can delete to go back to the standard blue, or specify others
#  theme(legend.key = element_blank(),  #removes the box around each legend item
#        legend.position = "bottom", #legend at the bottom
#        legend.direction = "horizontal",
#        legend.box = "horizontal",
#        legend.box.just = "centre")

```

```{r}
#ordisurf:
meta.nmds.mite <- metaMDS(NMDS.log)
ordi <- ordisurf(meta.nmds.mite ~ scrs_filt$folded_aspect) #created the ordisurf object
ordi.grid <- ordi$grid #extracts the ordisurf object
str(ordi.grid) #it's a list though - cannot be plotted as is
ordi.mite <- expand.grid(x = ordi.grid$x, y = ordi.grid$y) #get x and ys
ordi.mite$z <- as.vector(ordi.grid$z) #unravel the matrix for the z scores
ordi.mite.na <- data.frame(na.omit(ordi.mite)) #gets rid of the nas
ordi.mite.na #looks ready for plotting!
 
 
#data for plotting
##NMDS points
mite.NMDS.data <- scrs_filt #there are other ways of doing this. But this is the way I do it for ease of plotting
mite.NMDS.data$NMDS1 <- meta.nmds.mite$points[ ,1] #this puts the NMDS scores for the plots into a new dataframe. you could put them into an existing one if you preferred.
mite.NMDS.data$NMDS2 <- meta.nmds.mite$points[ ,2]
 
#setting up a plot in base graphics
plot(mite.NMDS.data$NMDS1, mite.NMDS.data$NMDS2,
     xlab="NMDS1", ylab="NMDS2",  
     pch = c(1,16)[as.factor(mite.NMDS.data$Topo)]) #symbols coded by factor
legend(title = "Topo", "bottomleft", levels(mite.NMDS.data$Topo), pch = c(1,16))
ordisurf(meta.nmds.mite~mite.env$SubsDens, col = "grey50", add = TRUE)
with(ordi.grid, contour(x = x, y = y, z = z, add=TRUE)) # you can use this as a check to make sure your ordigrid is doing the same thing as an ordisurf (ie, you've extracted the components correctly)
 

## Plotting in ggplot2
mite.ordisurf.ggplot<-ggplot(mite.NMDS.data, aes(x = NMDS1, y = NMDS2))+
  stat_contour(data = ordi.mite.na, aes(x = x, y = y, z = z, colour = rev("aspect")),
               binwidth = 2)+ #can change the binwidth depending on how many contours you want
  geom_point(size = 3, alpha = 0.8,  aes(colour = elev)) + #plots the NMDS points, with shape by topo type
  theme_bw() + #for aesthetics
  scale_shape_manual("Topo Type",  values = c(1,16)) + #sets the name of the legend for shape, and says which symbols we want (equivalent to the pch command in plot)
  labs(colour = "Substrate Density")+ #another way to set the labels, in this case, for the colour legend
  scale_colour_gradient(high = "darkgreen", low = "darkolivegreen1")+ #here we set the high and low of the colour scale.  Can delete to go back to the standard blue, or specify others
  theme(legend.key = element_blank(),  #removes the box around each legend item
        legend.position = "bottom", #legend at the bottom
        legend.direction = "horizontal",
        legend.box = "horizontal",
        legend.box.just = "centre")
mite.ordisurf.ggplot

?rev
```

```{r}
meta.nmds.mite <- metaMDS(mite) #no transformation of species data is made here prior to bray curtis dissimilarities being calculated. (Bray Curtis is the default in R).
meta.nmds.mite #prints very basic description
stressplot(meta.nmds.mite) # To gain the stress plot for stress values for your MDS
 
#ordisurf:
ordi <- ordisurf(meta.nmds.mite ~ mite.env$SubsDens) #created the ordisurf object
ordi.grid <- ordi$grid #extracts the ordisurf object
str(ordi.grid) #it's a list though - cannot be plotted as is
ordi.mite <- expand.grid(x = ordi.grid$x, y = ordi.grid$y) #get x and ys
ordi.mite$z <- as.vector(ordi.grid$z) #unravel the matrix for the z scores
ordi.mite.na <- data.frame(na.omit(ordi.mite)) #gets rid of the nas
ordi.mite.na #looks ready for plotting!
 
 
#data for plotting
##NMDS points
mite.NMDS.data <- mite.env #there are other ways of doing this. But this is the way I do it for ease of plotting
mite.NMDS.data$NMDS1 <- meta.nmds.mite$points[ ,1] #this puts the NMDS scores for the plots into a new dataframe. you could put them into an existing one if you preferred.
mite.NMDS.data$NMDS2 <- meta.nmds.mite$points[ ,2]
 
#setting up a plot in base graphics
plot(mite.NMDS.data$NMDS1, mite.NMDS.data$NMDS2,
     xlab="NMDS1", ylab="NMDS2",  
     pch = c(1,16)[as.factor(mite.NMDS.data$Topo)]) #symbols coded by factor
legend(title = "Topo", "bottomleft", levels(mite.NMDS.data$Topo), pch = c(1,16))
ordisurf(meta.nmds.mite~mite.env$SubsDens, col = "grey50", add = TRUE)
with(ordi.grid, contour(x = x, y = y, z = z, add=TRUE)) # you can use this as a check to make sure your ordigrid is doing the same thing as an ordisurf (ie, you've extracted the components correctly)
 
## Plotting in ggplot2
mite.ordisurf.ggplot<-ggplot(mite.NMDS.data, aes(x = NMDS1, y = NMDS2))+
  stat_contour(data = ordi.mite.na, aes(x = x, y = y, z = z, colour = rev(..level..)),
               binwidth = 2)+ #can change the binwidth depending on how many contours you want
  geom_point(size = 3, alpha = 0.8,  aes(shape = Topo)) + #plots the NMDS points, with shape by topo type
  theme_bw() + #for aesthetics
  scale_shape_manual("Topo Type",  values = c(1,16)) + #sets the name of the legend for shape, and says which symbols we want (equivalent to the pch command in plot)
  labs(colour = "Substrate Density")+ #another way to set the labels, in this case, for the colour legend
  scale_colour_gradient(high = "darkgreen", low = "darkolivegreen1")+ #here we set the high and low of the colour scale.  Can delete to go back to the standard blue, or specify others
  theme(legend.key = element_blank(),  #removes the box around each legend item
        legend.position = "bottom", #legend at the bottom
        legend.direction = "horizontal",
        legend.box = "horizontal",
        legend.box.just = "centre")
mite.ordisurf.ggplot
```


```{r eval = F, include = F}
#all spp
p <- ggplot(scrs, aes(x = NMDS1, y = NMDS2)) +
  geom_point(mapping = aes(x = NMDS1, y = NMDS2, colour = sev)) +
  coord_fixed() + ## need aspect ratio of 1!
  geom_segment(data = spp.scrs,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.25, "cm")), colour = "grey") +
  geom_text(data = spp.scrs, aes(x = NMDS1, y = NMDS2, label = Species),
            size = 3) 
p
```

The NMDS using all native plants suggests no difference between unburnt and LS, but HS is very different (figure 1). Douglas-fir (PSME) cover is much higher in LS/UN, whereas deerbrush (CEIN) cover is highest in the HS. Other species that are strongly associated with one habitat are investigated below (see indicator species analysis). Time-since-fire was not included in this analysis.   
  
Interestingly, the number of species in high severity and low severity is not statistically different on a plot-by-plot basis (see Fig. 2). However there are two important contrasts between HS and LS:     

*   the NMDS shows that the species composition is different even if species richness isn't (figure 1), and     

*   there are more species restricted to HS areas than species restricted to LS/UN areas (42 out of 139 species collected were only found in HS burns vs. only 16 species found in LS or UN areas but not in HS).

```{r, echo=FALSE,message = F,fig.keep='all',}
plant_dat$cover[plant_dat$cover > 0] <- 1 # convert to pa
plant_matrix2 <- spread(data = plant_dat, key = species, value = cover, fill = 0)

plant_spp_num <- left_join(plant_matrix2, site_data, by = "site_id") %>%
  mutate(num_spp = rowSums(dplyr::select(.,ABCO:WOFI)))

plant_spp_mean <- plant_spp_num %>% 
  group_by(sev) %>%
  summarize(mean_items = mean(num_spp))

hist <- ggplot(plant_spp_num, aes(x = sev, y = num_spp)) +
  geom_boxplot() +
  labs(title = "Figure 2: Number of species in each habitat")
hist


# or use plant_spp_mean <- tapply(plant_spp_num$num_spp, INDEX = plant_spp_num$sev, FUN = mean)

#str(plant_spp_num$sev)

#plant_nums <- aggregate(num_spp ~ sev, data = plant_spp_num, FUN = count)

#sepal.lengths <- tapply(plant_spp_num$num_spp, INDEX = iris$Species, FUN = mean)

hs <- plant_spp_num %>% filter(sev =="h") %>% dplyr::select(c(ABCO:WOFI))
hs1 <- rbind(hs, colSums(hs))
plant_dat$cover[plant_dat$cover > 0] <- 1 
hs1[hs1 > 0] <-1
hs2 <- cbind(hs1, rowSums(hs1)) #119 species in HS (23 sites)
#hs3 <-hs2 %>% dplyr::select(hs2[24,] > 0)
hs3 <- hs2[,(hs2[24,]) == 1]
hs_spp <- data.frame(names = colnames(hs3))

ls <- plant_spp_num %>% filter(sev =="l") %>% dplyr::select(c(ABCO:WOFI))
ls1 <- rbind(ls, colSums(ls))
ls1[ls1 > 0] <-1
ls2 <- cbind(ls1, rowSums(ls1)) #90 species in LS (19 sites)
ls3 <- ls2[,(ls2[20,]) == 1]
ls_spp <- data.frame(names = colnames(ls3))

un <- plant_spp_num %>% filter(sev =="u") %>% dplyr::select(c(ABCO:WOFI))
un1 <- rbind(un, colSums(un))
un1[un1 > 0] <-1
un2 <- cbind(un1, rowSums(un1)) #42 species in un (6 sites)
un3 <- un2[,(un2[7,]) == 1]
un_spp <- data.frame(names = colnames(un3))

#spp in HS not in LS
x1 <- cbind.fill(hs_spp, ls_spp, fill = NA)
names(x1)[1]<-"hs_spp1"
names(x1)[2]<-"ls_spp1"
hs_spp1 <- x1[,1]
ls_spp1 <- x1[,2]
#kable(spp_diff <- setdiff(hs_spp1, ls_spp1))

#spp in LS not in HS
x2 <- cbind.fill(ls_spp, hs_spp, un_spp, fill = NA)
#hs_spp1 <- x2[,1]
#ls_spp1 <- x2[,2]
#un_spp1 <- x2[,3]
#kable(spp_diff <- setdiff(hs_spp1, ls_spp1))

#spp in HS not in LS or UN
hs_spp1 <- rep(NA, 119) #change this number to the total number of species HS
ls_spp1 <- x2[,3]
un_df <- data.frame(hs_spp1, ls_spp1)
x3 <- rbind(x1, un_df)
hs_spp1 <- x3[,1]
ls_spp1 <- x3[,2]
spp_diff1 <- setdiff(hs_spp1, ls_spp1)

spp_diff1 <- data.frame(spp_diff1)
names(spp_diff1)[1]<-"species"
spp_diff1 <- left_join(spp_diff1, plant_names, by = "species") %>% dplyr::select(full_name)

#spp in LS or UN not in HS
spp_diff2 <- setdiff(ls_spp1, hs_spp1)
spp_diff2 <- data.frame(spp_diff2)
names(spp_diff2)[1] <- "species"
spp_diff2 <- left_join(spp_diff2, plant_names, by = "species") %>% 
  dplyr::select(full_name)

names(spp_diff1)[1]<-"Species found only in HS"
names(spp_diff2)[1]<-"Species found only in LS/UN"
spp_diff_table <- list(spp_diff1, spp_diff2) 

kable(spp_diff_table, caption = "Table 1: species restricted to only HS or LS/UN") %>%
  kable_styling()
```


####Birds

The NMDS for birds shows a very similar picture to the NMDS for plants (figure 2). The HS sites are clearly different from the LS and UN sites. There is a clear differentiation between species that favor one habitat or the other.

```{r, echo=FALSE,results='hide',fig.keep='all'}
# optional:remove birds with few sightings; not sure this makes much difference
## add columns true/false depending on obs count reaching minimum value
#bird_matrix <- rbind(bird_matrix1[,-c(1:20)], c("colsum", colSums(bird_matrix1[,-c(1:20)]) == 1))
bird_matrix <- bird_mat_w[,-c(1:37)] %>%
  bind_rows(summarise_all(., funs(sum(.)))) # note: if wanting to make a row named total, could have used: funs(if(is.numeric(.)) sum(.) else "Total"
  
## keep only columns where minimum is reached (if subsetting some sites, change row number, here 49)
bird_matrix <- bird_matrix[, (bird_matrix[111, ]) > 0]
## adding back in the site id and removing true/false row (and making sure it's all read in as numeric)
bird_matrix <- cbind(site_id = bird_mat_w[, 1], bird_matrix[-111, ]) %>%
  mutate_if(is.character, as.numeric)

# Make Bray-Curtis (?) dissimilarity matrix
#birds_matrix <- as.matrix(bird_matrix)

# calculate distance for NMDS.
NMDS.log<-bird_mat_w[,-1]#[,-c(1:20)])
sol <- metaMDS(NMDS.log)

scrs <- as.data.frame(scores(sol, display = "sites"))
scrs <- cbind(scrs, site_id = bird_mat_w$site_id)
scrs <- merge(scrs, site_data, by = "site_id")

vf <- envfit(sol, NMDS.log, perm = 9)

spp.scrs <- as.data.frame(scores(vf, display = "vectors"))
spp.scrs <- cbind(spp.scrs, Species = rownames(spp.scrs))

# To only show statistically significant species:
ef<-envfit(sol$points, NMDS.log, perm=1000)
ef.df<-as.data.frame(ef$vectors$arrows*sqrt(ef$vectors$r))
ef.df$species<-rownames(ef.df)

#only significant pvalues
#shortcutting ef$vectors
A <- as.list(ef$vectors)
#creating the dataframe
pvals<-as.data.frame(A$pvals)
arrows<-as.data.frame(A$arrows*sqrt(A$r))
C<-cbind(arrows, pvals)
#subset
Cred<-subset(C,pvals<0.01)
Cred <- cbind(Cred, Species = rownames(Cred))

#only stat. signif. spp
# maybe subset only species detected at least 5 times?
p1 <- ggplot(scrs, aes(x = NMDS1, y = NMDS2)) +
  geom_point(aes(colour = sev)) +
  coord_fixed() + ## need aspect ratio of 1!
  stat_ellipse(aes(colour = sev)) +
  geom_segment(data = Cred,
               aes(x = 0, xend = MDS1, y = 0, yend = MDS2),
               arrow = arrow(length = unit(0.25, "cm")), colour = "grey") +
  geom_text(data = Cred, 
            aes(x = MDS1, y = MDS2, label = Species),
            size = 3) +
  labs(title = "Figure 2: NMDS for birds")
p1

```

#### Lichens
```{r, echo=FALSE,results='hide',fig.keep='all'}

# calculate distance for NMDS.
# for species: 
#NMDS.log<-log1p(lichen_dat_wide[,-1])#[,-c(1:20)])
#for genus:
NMDS.log<-lichen_mat_species_pa_w %>% dplyr::select(-c(site_id)) #[,-c(1:20)])
           
sol <- metaMDS(NMDS.log)

scrs <- as.data.frame(scores(sol, display = "sites"))
scrs <- cbind(scrs, site_id = lichen_mat_species_w$site_id)
scrs <- merge(scrs, site_data, by = "site_id")

vf <- envfit(sol, NMDS.log, perm = 9)

spp.scrs <- as.data.frame(scores(vf, display = "vectors"))
spp.scrs <- cbind(spp.scrs, Species = rownames(spp.scrs))

# To only show statistically significant species:
ef<-envfit(sol$points, NMDS.log, perm=1000)
ef.df<-as.data.frame(ef$vectors$arrows*sqrt(ef$vectors$r))
ef.df$species<-rownames(ef.df)

#only significant pvalues
#shortcutting ef$vectors
A <- as.list(ef$vectors)
#creating the dataframe
pvals<-as.data.frame(A$pvals)
arrows<-as.data.frame(A$arrows*sqrt(A$r))
C<-cbind(arrows, pvals)
#subset
Cred<-subset(C,pvals<0.001)
Cred <- cbind(Cred, Species = rownames(Cred))

#only stat. signif. spp
p <- ggplot(scrs) +
  geom_point(mapping = aes(x = NMDS1, y = NMDS2, colour = sev)) +
  coord_fixed() + ## need aspect ratio of 1!
  stat_ellipse(level = 0.9, aes(x = NMDS1, y = NMDS2, colour = sev)) +
  geom_segment(data = Cred,
               aes(x = 0, xend = MDS1, y = 0, yend = MDS2),
               arrow = arrow(length = unit(0.25, "cm")), colour = "grey") +
  geom_text(data = Cred, aes(x = MDS1, y = MDS2, label = Species),
            size = 3) +
  labs(title = "Figure 3: NMDS for lichens")
p
```


#### Plants
```{r, echo=FALSE,results='hide',fig.keep='all'}
# calculate distance for NMDS
NMDS.log<-plant_mat_pa_w[,-1]#[,-c(1:20)])
sol <- metaMDS(NMDS.log)

scrs <- as.data.frame(scores(sol, display = "sites"))
scrs <- cbind(scrs, site_id = plant_mat_pa_w$site_id)
scrs <- merge(scrs, site_data, by = "site_id")

vf <- envfit(sol, NMDS.log, perm = 9)

spp.scrs <- as.data.frame(scores(vf, display = "vectors"))
spp.scrs <- cbind(spp.scrs, Species = rownames(spp.scrs))

# To only show statistically significant species:
ef<-envfit(sol$points, NMDS.log, perm=1000)
ef.df<-as.data.frame(ef$vectors$arrows*sqrt(ef$vectors$r))
ef.df$species<-rownames(ef.df)

#only significant pvalues
#shortcutting ef$vectors
A <- as.list(ef$vectors)
#creating the dataframe
pvals<-as.data.frame(A$pvals)
arrows<-as.data.frame(A$arrows*sqrt(A$r))
C<-cbind(arrows, pvals)
#subset
Cred<-subset(C,pvals<0.05)
Cred <- cbind(Cred, Species = rownames(Cred))

#only stat. signif. spp
p <- ggplot(scrs, mapping = aes(x = NMDS1, y = NMDS2, colour = sev)) +
  geom_point() +
  coord_fixed() + ## need aspect ratio of 1!
  geom_segment(data = Cred,
               aes(x = 0, xend = MDS1, y = 0, yend = MDS2),
               arrow = arrow(length = unit(0.25, "cm")), colour = "grey") +
  geom_text(data = Cred, aes(x = MDS1, y = MDS2, label = Species),
            size = 3) +
  labs(title = "Figure 4: NMDS for plants")
p
```

#### All taxa
```{r, echo=FALSE,results='hide',fig.keep='all'}
# read in data
#plant_dat_long <- plant_dat %>% 
#  mutate(pa = cover) %>% 
#  dplyr::select(site_id, species, pa)

#lichen_dat_genus_pa_l <- lichen_dat_genus_pa_l %>% 
#  mutate(species = genus) %>% 
#  dplyr::select(site_id, species, pa)

#insect_dat_long_filt1 <- insect_dat_l %>% 
#  dplyr::select(-c(2:28)) %>% 
#  filter(number != 0)

#insect_dat_long_filt <- splist2presabs(insect_dat_long_filt1, sites.col = 1, sp.col = 2)  %>% 
#  gather(key = species, value = pa, Aculeata:Thysanoptera)


# calculate distance for NMDS.
# for species: 
#NMDS.log<-log1p(lichen_dat_wide[,-1])#[,-c(1:20)])
#for genus:
NMDS.log<-(all_spp_w_pa[,-1])#[,-c(1:20)])
sol <- metaMDS(NMDS.log, k = 2)

scrs <- as.data.frame(scores(sol, display = "sites"))
scrs <- cbind(scrs, site_id = all_spp_w_pa$site_id)
scrs <- merge(scrs, site_data, by = "site_id")

vf <- envfit(sol, NMDS.log, perm = 9)

spp.scrs <- as.data.frame(scores(vf, display = "vectors"))
spp.scrs <- cbind(spp.scrs, Species = rownames(spp.scrs))

# To only show statistically significant species:
ef<-envfit(sol$points, NMDS.log, perm=1000)
ef.df<-as.data.frame(ef$vectors$arrows*sqrt(ef$vectors$r))

ef.df<-data.frame(ef$vectors$arrows,R = ef$vectors$r, P = ef$vectors$pvals)

ef.df$species<-rownames(ef.df)

#only significant pvalues
#shortcutting ef$vectors
A <- as.list(ef$vectors)
#creating the dataframe
pvals<-as.data.frame(A$pvals)
arrows<-as.data.frame(A$arrows*sqrt(A$r))
C<-cbind(arrows, pvals)
#subset
Cred<-subset(C,pvals<0.001)
Cred <- cbind(Cred, Species = rownames(Cred))

scrs$cov_cat <- as.factor(scrs$cov_cat)
#only stat. signif. spp
p <- ggplot(scrs, mapping = aes(x = NMDS1, y = NMDS2)) +
  geom_point(aes(colour = sev)) +
  coord_fixed() + ## need aspect ratio of 1!
  stat_ellipse(level = 0.9, aes(colour = sev)) +
  geom_segment(data = Cred,
               aes(x = 0, xend = MDS1, y = 0, yend = MDS2),
               arrow = arrow(length = unit(0.25, "cm")), colour = "grey") +
  geom_text(data = Cred, aes(x = MDS1, y = MDS2, label = Species),
            size = 3) +
  labs(title = "Figure 4: NMDS for all spp")
p

ggplotly(p, tooltip = "site_id")

?ggplotly
?count

all_spp %>% count(species, sort = TRUE)

species
species %>% count(species, sort = TRUE)

p3d <- plot_ly(scrs, x = ~NMDS1, y = ~NMDS2, z = ~NMDS3, color = ~sev, symbol = ~sev, symbols = (c("square", "diamond", "circle", "cross", "x")), text = ~paste('cc: ', tree_cov)) %>%
  add_markers() %>%
  layout(scene = list(xaxis = list(title = 'NMDS1'),
                     yaxis = list(title = 'NMDS2'),
                     zaxis = list(title = 'NMDS3')))
p3d


```

### Richness and diversity

```{r}
# Number of occurrences of all species
all_spp_dat_l %>% 
  group_by(species) %>% 
  summarise(sums=sum(pa))

#same thing, per sev type
all_spp_dat_l %>%
  filter(pa != 0) %>% 
  dplyr::count(species, sev, sort = TRUE)

df <- all_spp_dat_l %>%
#  filter(pa != 0) %>% 
  add_count(species)


# species richness at each plot
# from http://www.flutterbys.com.au/stats/tut/tut13.2.html
df <- apply(all_spp_w_pa[,-1]>0,1,sum)

#Same but in data frame format
all_spp_richness <- ddply(all_spp_w_pa,~site_id,function(x) {
   data.frame(richness=sum(x[-1]>0))
 })

# Shannon diversity
diversity(all_spp_w_pa[-1], index="shannon")

#OR
all_spp_shannon <- ddply(all_spp_w_pa,~site_id,function(x) {
        data.frame(shannon=diversity(x[-1], index="shannon"))
})

#Simpson diversity
diversity(all_spp_w_pa[-1], index="simpson")

#OR
all_spp_simpson <- ddply(all_spp_w_pa,~site_id,function(x) {
        data.frame(simpson=diversity(x[-1], index="simpson"))
})

#put it together
all_spp_div <- merge(all_spp_richness, all_spp_shannon, by = "site_id") %>%
  merge(., all_spp_simpson, by = "site_id") %>% 
  merge(., site_data, by = "site_id")

#all_spp_div  %>%
#  group_by(sev) %>%
#  summarise_all(mean, na.rm=TRUE)

all_spp_div$cov_cat <- as.factor(all_spp_div$cov_cat)

ggplot(all_spp_div, aes(x = cov_cat, y = richness)) +
  geom_boxplot()

ggplot(all_spp_div, aes(x = cov_cat, y = shannon)) +
  geom_boxplot()

ggplot(all_spp_div, aes(x = cov_cat, y = simpson)) +
  geom_boxplot()

# do anovas to test differences?
```


### Beta diversity (plants)

```{r}
# Code from Jonah

#all spp
comptable <-  all_spp_w_pa %>% full_join(site_data[, c("site_id", "sev")], by = "site_id") %>% 
  mutate(sev3 = ifelse(sev == "u" | sev == "l", "lu", sev)) %>% dplyr::select(-c(site_id, sev))


#birds
comptable <-  bird_dat_w %>% 
  mutate(sev3 = ifelse(sev == "u" | sev == "l", "lu", sev)) %>% 
  dplyr::select(c(sev3, ACWO:YRWA)) %>% 
  dplyr::select(-V1) #puttig together the plot id and sev and joinign wide format data then dropping site id

#plants
comptable <-  plant_dat_pa_w %>% 
  mutate(sev3 = ifelse(sev == "u" | sev == "l", "lu", sev)) %>% 
  dplyr::select(c(sev3, ABCO:YAMI)) 
  

betajaccard <- function(df){
  
  tempdf = df %>% dplyr::select(-sev3)
  
  dis = c(vegdist(tempdf, method = "binomial", binary = "true", na.rm = TRUE))
  
  result = data.frame(sev3 = rep(unique(df$sev3), length = length(dis)),
#                      year = rep(unique(df$year), length = length(dis)),
                      dis = dis)
  
}

bj_output = comptable %>% group_by(sev3) %>% do(betajaccard(.))

?vegdist
d <- betadiver(comptable, "w")

bj_output %>% ggplot(aes(x = sev3,
                         y = dis,
                         fill = sev3,
                         group = sev3)) +
  geom_boxplot() +
  #facet_wrap(~year) +
  xlab("Fire Severity") +
  ylab("beta div") +
#  ylim(0, 1) +
  #scale_fill_viridis_c(option = "magma") +
  ggtitle("Pairwise Beta Diversity (binomial)")



#???
sampdat = comp.matrix %>% filter(Regen_ID %in% c("ANG001", "ANG005") & year == 2008)
t(sampdat)
vegdist(sampdat[,-c(1,2)],
        method = "jaccard", binary = "true")
```

10/19: Using the canopy cover categories, there is a significant impact of groups on plants (using %cov). It looks like the highest cover has the lowest diversity. But using all spp, beta div (distance to centroid) is higher! Totally flat for birds. BUT what is this calculating???


```{r}
all_spp_dat_l <- gather(all_spp_dat_w, key = site_id, value = pa, Ahtiana:YAMI)

# plot from https://www.fromthebottomoftheheap.net/2016/04/17/new-plot-default-for-betadisper/
#install.packages("devtools")
#devtools::install_github("vegandevs/vegan")
#args(vegan:::plot.betadisper)


plant_glm_cov2 <- plant_glm_cov %>% 
  mutate(sev_tsf2 = paste(sev, tsf_cat2))

# can use betapart package here: dist<-beta.pair(plant_matrix1[,-1], index.family="jaccard") FOR PRESENCE ABSENCE
# for abundance data, use dist <- bray.part(plant_matrix1[,-1])
# In both cases use dist[[3]] instead of dis in betadisper


#all spp
dis <- vegdist(all_spp_w_pa[,-c(1)], method = "jaccard", binary = T)
groups <- factor(all_spp_dat_w$sev)

#plants
dis <- vegdist(plant_mat_cov_w[,-c(1,183)], method = "jaccard", binary = T) 
groups <- factor(plant_dat_pa_w$cov_cat)

#birds
#dis <- vegdist(bird_mat_w[,(38:104)]) 
dis <- vegdist(bird_mat_w[,(38:104)], method = "jaccard", binary = T) 
groups <- factor(bird_dat_w$cov_cat)

mod <- betadisper(dis, groups)  # Calculate multivariate dispersions

scores(mod)

plot(mod)
boxplot(mod)
anova(mod)









mod1 <-melt(mod)

df <- distance(all_spp_w_pa[,-1], method = "jaccard")

spe.dis <- vegdist(log1p(spe), method='bray')

# do land use categories differ in community composition? clustering of sites using dissimilarity matrix & UPGMA clustering

library(RCurl)
library(vegan)
library(gclus)
library(RColorBrewer)
library(ade4)
library(adespatial)

all_sppp <-all_spp_w_pa[,-c(1)]


kib <-read.csv(text=getURL("https://raw.githubusercontent.com/VFugere/QCBS_betadiv/master/kibale_inverts.csv"), header=T)
colnames(kib)[1] <- 'site.code' #corrects excel generated junk
kib$land.use <- relevel(kib$land.use, 'park') #more intuitive to have park/forest as reference level
#have a look at the data
head(kib)
str(kib)
#some useful matrices
spe <- kib[,-c(1:3)]
row.names(spe) <- kib$site.code
site <- kib[,2:3]
rownames(site) <- kib[,1]


spe.dis.UPGMA <- hclust(dis, method="average")
spe.dis.o <- reorder.hclust(spe.dis.UPGMA, dis)

dend <- as.dendrogram(spe.dis.o)
com.tx <- apply(t(all_spp_w_pa[,-c(1)]),1,sum) #common taxa for heat map plot
spe.com <- all_sppp[,which(com.tx > 20)]
or <- vegemite(spe.com, spe.dis.o,scale='log')
heatmap(t(spe.com[or$species]), Rowv=NA, Colv=dend, cexRow=0.5, cexCol=0.5, col=c('white', brewer.pal(5,"Greens")), scale="column", revC = T, margin=c(4.5,8))

#beta diversity, at last
beta.BC <- betadisper(dis, groups, type = 'centroid')
plot(beta.BC) #ugly

mybetadivplot <- function (x, cex=1){
  g <- scores(x, choices = 1:2)
  plot(g$sites, type = "n", ylab='Dimension 2', xlab='Dimension 1')
  points(g$sites, pch = c(21,22,21)[all_spp_dat_w$cov_cat], bg=c('white','white','black')[all_spp_dat_w$cov_cat])
  points(g$centroids, pch = 16, cex = 1.5, col = "red")
  for (i in levels(x$group)) {
    ch <- chull(g$sites[x$group == i, ])
    ch <- c(ch, ch[1])
    lines(x$vectors[x$group == i, 1:2][ch, ], col = "black", lty = "dashed")
  }
#  legend('topright',cex=1, pt.cex=1, inset = c(0.2,0),legend=c("Kab (farm)","Kam (farm)",'For (park)'),pch=c(21,22,21),pt.bg=c('white','white','black'),bty='n')
}
mybetadivplot(beta.BC)
legend('bottomright',bty='n',inset=c(0.2,0),legend = expression(italic(Bray-Curtis)))










  # remove self-comparisons
  df = df %>%
    filter(as.character(Var1) != as.character(Var2)) %>%
    mutate_if(is.factor,as.character)
  
  # get sample data (S4 error OK and expected)
  sd = sample_data(p) %>%
    select(s, d) %>%
    mutate_if(is.factor,as.character)
  
  # combined distances with sample data
  colnames(sd) = c("Var1", "Type1")
  wu.sd = left_join(wu.m, sd, by = "Var1")
  
  colnames(sd) = c("Var2", "Type2")
  wu.sd = left_join(wu.sd, sd, by = "Var2")
  
  # plot
  ggplot(wu.sd, aes(x = Type2, y = value)) +
    theme_bw() +
    geom_point() +
    geom_boxplot(aes(color = ifelse(Type1 == Type2, "red", "black"))) +
    scale_color_identity() +
    facet_wrap(~ Type1, scales = "free_x") +
    theme(axis.text.x=element_text(angle = 90, hjust = 1, vjust = 0.5)) + 
    ggtitle(paste0("Distance Metric = ", m))
}


# set binary: TRUE if presence/absence data
#             FALSE if abundance data
binary <- TRUE

m <- 500
n <- m
set.seed(12345)
if(binary == TRUE){
  # Create matrix with random 0/1 values
  x <- sample.int (2, m*n, TRUE)-1L
  dim(x) <- c(m,n)  
} else {
  # Or create matrix with values in the range 0-10
  x <- matrix(round(runif(n*m, 0, 10)), nrow = m, ncol = n)  
}

#### Test the standardization 
jaccard <- vegdist(x, method = "jaccard", binary = FALSE)
jaccard.stand <- vegdist(x, method = "jaccard", binary = TRUE)
### Are they equal?
all(jaccard == jaccard.stand)

### Plot for safety
jacc.df <- melt(cbind(jaccard, jaccard.stand))
ggplot(jacc.df, aes(x = value, fill = Var2, colour = Var2)) + 
  geom_histogram(alpha = 0.2, position = "identity", binwidth = 0.01)

str(jaccard)

```

```{r}
# from ?oecosimu

## Use the first eigenvalue of correspondence analysis as an index
## of structure: a model for making your own functions.
data(sipoo)
out <- oecosimu(sipoo, decorana, "swap", burnin=100, thin=10, statistic="evals")
out
## Inspect the swap sequence as a time series object
plot(as.ts(out))
lag.plot(as.ts(out))
acf(as.ts(out))
## Density plot: needs lattice
require(lattice)

densityplot(out, as.table = TRUE)
## Use quantitative null models to compare
## mean Bray-Curtis dissimilarities
data(dune)
meandist <- function(x) mean(vegdist(x, "bray"))
mbc1 <- oecosimu(dune, meandist, "r2dtable")
mbc1
## Define a custom function that shuffles
## cells in each row
f <- function(x) {
    apply(x, 2, function(z) sample(z, length(z)))
}
mbc2 <- oecosimu(as.matrix(dune), meandist, f)
mbc2
```


### Plots for plants

```{r}
plant_plot <- plant_dat_pa_w  %>% 
  mutate(num_spp = rowSums(dplyr::select(.,ACMI:WOFI)))

# Spp number in each habitat
ggplot(plant_plot, aes(x = sev_tsf, y = num_spp)) +
  geom_boxplot() +
  labs(title = "Figure X: spp num in each habitat", y = "# spp", x = "habitat")

ggplot(plant_plot, aes(x = tsf, y = num_spp, color = sev)) +
  geom_smooth(method = "loess", aes(group = sev)) +
  geom_jitter() +
  labs(title = "Figure X: spp num in each habitat", y = "number of spp", x = "TSF")

ggplot(plant_plot, aes(x = sev_tsf, y = num_spp, color = sev)) +
  geom_boxplot(aes(group = sev_tsf)) +
  geom_jitter() +
  labs(title = "Figure X: spp num in each habitat", y = "number of spp", x = "TSF")

?geom_boxplot


# Relationship of spp number to shrub cover
ggplot(plant_plot, aes(x = shrub_cov, y = num_spp)) +
  geom_jitter(aes(color = sev)) +
  geom_smooth(method = "glm") +
  labs(title = "Figure X: shrub cover vs spp number", y = "# spp", x = "shrub cover (%)")

# Relationship of spp number to tree cover
ggplot(plant_plot, aes(x = tree_cov, y = num_spp)) +
  geom_jitter(aes(color = sev_tsf2)) +
  geom_smooth(method = "glm") +
  labs(title = "Figure X: tree cover vs spp number", y = "# spp", x = "tree cover (%)")

# Relationship of spp number to herbaceous cover
ggplot(plant_plot, aes(x = herb_cov, y = num_spp)) +
  geom_jitter(aes(color = sev)) +
  geom_smooth(method = "glm") +
  labs(title = "Figure X: herbaceous cover vs spp number", y = "# spp", x = "herbaceous cover (%)")

# summarize data
data_summary <- function(data, varname, groupnames){
  require(plyr)
  summary_func <- function(x, col){
    c(mean = mean(x[[col]], na.rm=TRUE),
      sd = sd(x[[col]], na.rm=TRUE))
  }
  data_sum<-ddply(data, groupnames, .fun=summary_func,
                  varname)
  data_sum <- rename(data_sum, c("mean" = varname))
 return(data_sum)
}

df <- data_summary(plant_plot, varname="num_spp", groupnames="sev_tsf2")
df

#plot
hist <- ggplot(df, aes(x = tsf_cat2, y = num_spp)) +
  geom_histogram(stat = "identity") +
  labs(title = "Figure X: number of species in each habitat", y = "# spp", x = "habitat") +
  geom_errorbar(aes(ymin=num_spp-sd, ymax=num_spp+ sd), width=.2) 
hist

#%>%
#  group_by(sev, tsf) %>%
#  summarize(mean_size = mean(num_spp, na.rm = TRUE))

ggplot(plant_plot, aes(x = num_spp)) +
  geom_bar(aes(colour = sev))

ggplot(plant_mrpp_d, aes(x = sev_tsf, y = tree_cov)) +
  geom_boxplot()

ggplot(plant_mrpp_d, aes(x = sev, y = shrub_cov)) +
  geom_boxplot()

ggplot(plant_mrpp_d, aes(x = sev, y = herb_cov)) +
  geom_boxplot()

ggplot(plant_mrpp_d, aes(x = sev, y = bare_cov)) +
  geom_boxplot()

# tree cover and bare ground are positively correlated
ggplot(plant_mrpp_d, aes(x = tree_cov, y = bare_cov)) +
  geom_point(aes(colour = sev))+
  geom_smooth(method ="lm")

# no relationship between herb cover and tree cover
ggplot(plant_mrpp_d, aes(x = tree_cov, y = herb_cov)) +
  geom_point(aes(colour = sev))+
  geom_smooth(method ="lm")

# shrub cover and tree cover are negatively correlated
ggplot(plant_mrpp_d, aes(x = tree_cov, y = shrub_cov)) +
  geom_point(aes(colour = sev))+
  geom_smooth(method ="lm")

# tree cover increases with tsf in HS
ggplot(plant_mrpp_d, aes(x = tsf, y = tree_cov, group = sev)) +
  geom_point(aes(colour = sev))+
  geom_smooth(method ="lm", aes(colour = sev))

# bare ground decreased with tsf in LS, increases with tsf in HS (although this is less significant)
ggplot(plant_mrpp_d, aes(x = tsf, y = bare_cov, group = sev)) +
  geom_point(aes(colour = sev))+
  geom_smooth(method ="lm", aes(colour = sev))

ggplot(plant_spp_num, aes(x = tsf, y = num_spp, group = sev)) +
  geom_point(aes(colour = sev))+
  geom_smooth(method ="lm", aes(colour = sev))

ggplot(plant_spp_num, aes(x = tree_cov, y = num_spp)) +
  geom_point(aes(colour = sev))+
  geom_smooth(method ="lm")

ggplot(plant_spp_num, aes(x = shrub_cov, y = num_spp)) +
  geom_point(aes(colour = sev))+
  geom_smooth(method ="lm")

ggplot(plant_spp_num, aes(x = bare_cov, y = num_spp)) +
  geom_point(aes(colour = sev))+
  geom_smooth(method ="lm")

p <- plant_spp_num %>% 
  dplyr::select(site_id, num_spp) %>% 
  mutate_if()
                
scrs1 <- left_join(scrs, p, by = "site_id")

ggplot(scrs1, aes(x = herb_cov, y = NMDS1)) +
  geom_point(aes(colour = sev))+
  geom_smooth(method ="lm")

#Need to look at snag density/BA, total live tree density/BA, max tree height, proportion of conifer to hardwood cover



```
### Plots for birds

```{r}
bird_mat_num <- bird_matrix1 %>%
  mutate(num_spp = rowSums(dplyr::select(.,ACWO:YEWA)))

#doesn't look right:
hist <- ggplot(bird_mat_num, aes(x = sev, y = num_spp)) +
  geom_boxplot() +
  labs(title = "Figure 2: Number of bird species in each habitat")
hist

#the following two should be split between HS and LS/UN rather than showing all together
ggplot(bird_mat_num, aes(x = tree_cov, y = num_spp)) +
  geom_point(aes(colour = sev))+
  geom_smooth(method ="lm")

ggplot(bird_mat_num, aes(x = shrub_cov, y = num_spp)) +
  geom_point(aes(colour = sev))+
  geom_smooth(method ="lm")
```

### Plots for lichens
```{r}
# Bar graph with species frequency

lichen_totals <- read.csv("data/lichen_data.csv", header = T) %>% 
  dplyr::select(site_id, species, abund) %>% 
  spread(key = species, value = abund, fill = 0) %>%
  apply(., 2, function(c)sum(c!=0))

lichen_dat_totals1 <- lichen_mat_species_w[,-1] %>% 
  bind_rows(., lichen_totals)

lichen_dat_totals <- lichen_dat_totals1[-c(1:48),-60] %>% 
  gather(key = species, value = frequency) %>% 
  arrange(desc(frequency)) %>% 
  filter(frequency > 11)

ggplot(lichen_dat_totals, aes(x = reorder(species, -frequency), y = frequency)) +
         geom_col()

# Lichens of HS
lichen_HS <- read.csv("data/lichen_data.csv", header = T) %>%
  mutate(site_id = as.factor(site_id)) %>% 
  dplyr::select(site_id, species, abund) %>% 
  left_join(site_data, by = "site_id") %>% 
  filter(sev == "h") %>% 
  filter(abund != 0) %>% 
  filter(!site_id %in% c("7", "6", "34", "60", "14", "36", "5", "28", "16")) %>% 
  spread(key = species, value = abund, fill = 0)

lichen_HS_list1 <- lichen_HS %>% 
  dplyr::select(-c(1:30)) %>% 
  dplyr::select(-c(CONI, PECO, FUME, NOPU))

lichen_HS_list2 <- t(lichen_HS_list1)

lichen_HS_list2 <- cbind(rownames(lichen_HS_list2), data.frame(lichen_HS_list2, row.names=NULL)) 

names(lichen_HS_list2)[1]<-"species"

lichen_HS_list <- lichen_HS_list2 %>% 
  left_join(., lichen_names, by = "species") %>% 
  dplyr::select(full_name)

#write.table(lichen_HS_list, file = "lichen_HS_list.csv")







plant_dat$cover[plant_dat$cover > 0] <- 1 # convert to pa
plant_matrix2 <- spread(data = plant_dat, key = species, value = cover, fill = 0)

plant_spp_num <- left_join(lichen_mat_genus_pa_w, site_data, by = "site_id") %>%
  mutate(num_spp = rowSums(dplyr::select(.,Ahtiana:Xanthomendoza)))

plant_spp_mean <- plant_spp_num %>% 
  group_by(sev) %>%
  summarize(mean_items = mean(num_spp))

hist <- ggplot(plant_spp_num, aes(x = sev, y = num_spp)) +
  geom_boxplot() +
  labs(title = "Figure 2: Number of species in each habitat")
hist


# or use plant_spp_mean <- tapply(plant_spp_num$num_spp, INDEX = plant_spp_num$sev, FUN = mean)

#str(plant_spp_num$sev)

#plant_nums <- aggregate(num_spp ~ sev, data = plant_spp_num, FUN = count)

#sepal.lengths <- tapply(plant_spp_num$num_spp, INDEX = iris$Species, FUN = mean)

hs <- lichen_dat_genus_pa_w %>% filter(sev =="h") %>% dplyr::select(c(Ahtiana:Xanthomendoza))
hs1 <- rbind(hs, colSums(hs))
#plant_dat$cover[plant_dat$cover > 0] <- 1 
hs1[hs1 > 0] <-1
hs2 <- cbind(hs1, rowSums(hs1)) #121 species in HS (23 sites)
#hs3 <-hs2 %>% dplyr::select(hs2[24,] > 0)
hs3 <- hs2[,(hs2[24,]) == 1]
hs_spp <- data.frame(names = colnames(hs3))

ls <- lichen_dat_genus_pa_w %>% filter(sev =="l") %>% dplyr::select(c(Ahtiana:Xanthomendoza))
ls1 <- rbind(ls, colSums(ls))
ls1[ls1 > 0] <-1
ls2 <- cbind(ls1, rowSums(ls1)) #90 species in LS (19 sites)
ls3 <- ls2[,(ls2[20,]) == 1]
ls_spp <- data.frame(names = colnames(ls3))

un <- lichen_dat_genus_pa_w %>% filter(sev =="u") %>% dplyr::select(c(Ahtiana:Xanthomendoza))
un1 <- rbind(un, colSums(un))
un1[un1 > 0] <-1
un2 <- cbind(un1, rowSums(un1)) #42 species in un (6 sites)
un3 <- un2[,(un2[7,]) == 1]
un_spp <- data.frame(names = colnames(un3))

#spp in HS not in LS
x1 <- cbind.fill(hs_spp, ls_spp, fill = NA)
names(x1)[1]<-"hs_spp1"
names(x1)[2]<-"ls_spp1"
hs_spp1 <- x1[,1]
ls_spp1 <- x1[,2]
#kable(spp_diff <- setdiff(hs_spp1, ls_spp1))


#spp in HS not in LS or UN
hs_spp1 <- rep(NA, 26)
ls_spp1 <- x2[,3]
un_df <- data.frame(hs_spp1, ls_spp1)
x3 <- rbind(x1, un_df)
hs_spp1 <- x3[,1]
ls_spp1 <- x3[,2]
spp_diff1 <- setdiff(hs_spp1, ls_spp1)
spp_diff1 <- data.frame(spp_diff1)
names(spp_diff1)[1]<-"species"
spp_diff1 <- left_join(spp_diff1, plant_names, by = "species") %>% dplyr::select(species)

#spp in LS not in HS
x2 <- cbind.fill(ls_spp, hs_spp, un_spp, fill = NA)

#spp in LS or UN not in HS
spp_diff2 <- setdiff(ls_spp1, hs_spp1)
spp_diff2 <- data.frame(spp_diff2)
names(spp_diff2)[1] <- "species"
spp_diff2 <- left_join(spp_diff2, plant_names, by = "species") %>% 
  dplyr::select(full_name)

names(spp_diff1)[1]<-"Species found only in HS"
names(spp_diff2)[1]<-"Species found only in LS/UN"
spp_diff_table <- list(spp_diff1, spp_diff2) 

kable(spp_diff_table, caption = "Table 1: species restricted to only HS or LS/UN") %>%
  kable_styling()
```

### MRPP (plants)
```{r echo = F}
mrpp_data <- plant_mat_pa_w[,-1]
  
  
  plant_dat_cov_w %>% # [,c(2:82)]
  filter(sev %in% c("l","h", "u")) %>% 
  select(ABPR:YAMI) %>% 
  select(which(colSums(.) > 0)) 

mrpp_cat <- plant_dat_cov_w
  

mrpp_data <- div$rich_plant
mrpp_cat <- div
  
  
  plant_dat_cov_w %>% 
  filter(sev %in% c("l","h", "u"))
  
plant.mrpp <- mrpp(mrpp_data, mrpp_cat$sev) # 2:181 for all plants; 2:57 for trees/shrubs

plant.mrpp <- mrpp(plant_mat_cov_w[,-1], plant_dat_cov_w$sev)

# Save and change plotting parameters
def.par <- par(no.readonly = TRUE)
layout(matrix(1:2,nr=1))

#plot(plant.ord <- metaMDS(mrpp_data, type="text", display="sites"))
#ordihull(plant.ord, mrpp_cat$sev, label = T)

with(plant.mrpp, {
  fig.dist <- hist(boot.deltas, xlim=range(c(delta,boot.deltas)), 
                 main="Test of Differences Among Groups")
  abline(v=delta); 
  text(delta, 2*mean(fig.dist$counts), adj = -0.5,
     expression(bold(delta)), cex=1.5 )  }
)
par(def.par)

## meandist 
# calculates a matrix of mean within-cluster dissimilarities (diagonal) and between-cluster dissimilarities (off-diagonal elements)
plant.md <- meandist(vegdist(mrpp_data), mrpp_cat$sev) # Need to adjust the columns to only keep species.... should fix this
plant.md
summary(plant.md)
plot(plant.md)


out1<- aovp(rich_plant ~ sev, div)

anova(out1) # permutational anova

TukeyHSD(out1) # pairwise comparisons



?aovp


#test to see if any columns are not numeric
#plant_mat_cov_w[,-1] %>% select_if(negate(is.numeric))
```

## MRPP (insects) 
We see from the MRPP that the first 2 age categories for HS stand out, but the older HS burns cluster with the LS and UN.
```{r echo = F}
insect_dat_w <- insect_dat_w %>% 
  mutate(sev = ifelse(sev == "multiple", "h", sev))


insect.mrpp <- mrpp(insect_dat_w[,c(37:53)], insect_dat_w$sev, distance = "bray")

# Save and change plotting parameters
def.par <- par(no.readonly = TRUE)
layout(matrix(1:2,nr=1))

plot(insect.ord <- metaMDS(sqrt(insect_dat_w[,c(37:53)])), type="text", display="sites" )
ordihull(insect.ord, insect_dat_w$sev, label = T)

with(insect.mrpp, {
  fig.dist <- hist(boot.deltas, xlim=range(c(delta,boot.deltas)), 
                 main="Test of Differences Among Groups")
  abline(v=delta); 
  text(delta, 2*mean(fig.dist$counts), adj = -0.5,
     expression(bold(delta)), cex=1.5 )  }
)
par(def.par)

## meandist 
# calculates a matrix of mean within-cluster dissimilarities (diagonal) and between-cluster dissimilarities (off-diagonal elements)
insect.md <- meandist(vegdist(insect_dat_w[,c(37:53)]), insect_dat_w[["sev"]])
insect.md
summary(insect.md)
plot(insect.md)
```

## MRPP (birds) 

```{r echo = F}
bird_dat_w1 <- div %>% dplyr::select(sev, rich_bird)#%>% 
#  filter(yr == 2018) %>% 
#  mutate(sev = ifelse(sev == "multiple", "h", sev))

mrpp_data <- bird_dat_w1 %>% 
  dplyr::select(rich_bird)

bird.mrpp <- mrpp(mrpp_data, bird_dat_w1$sev)

?mrpp

bird.mrpp$Pvalue

# Save and change plotting parameters
def.par <- par(no.readonly = TRUE)
layout(matrix(1:2,nr=1))

plot(bird.ord <- metaMDS(mrpp_data, type="text", display="sites" ))
ordihull(bird.ord, bird_dat_w1$sev, label = T)

with(bird.mrpp, {
  fig.dist <- hist(boot.deltas, xlim=range(c(delta,boot.deltas)), 
                 main="Test of Differences Among Groups")
  abline(v=delta); 
  text(delta, 2*mean(fig.dist$counts), adj = -0.5,
     expression(bold(delta)), cex=1.5 )  }
)
par(def.par)

## meandist 
# calculates a matrix of mean within-cluster dissimilarities (diagonal) and between-cluster dissimilarities (off-diagonal elements)
bird.md <- meandist(vegdist(mrpp_data), bird_dat_w1$sev, method = "jaccard")
bird.md
summary(bird.md)
plot(bird.md)

str(bird.md)
bird.md
```

## MRPP (lichens) 

```{r echo = F}
#meandist not working now for some reason, same with plants above

lich_dat_mrpp <- lichen_dat_species_pa_w %>%dplyr::filter(sev == "Rx" | sev == "h")

mrpp_data <- lich_dat_mrpp %>% dplyr::select(-c(fire:tsf_num, site_id))

mrpp_cat <- lich_dat_mrpp

lichen.mrpp <- mrpp(mrpp_data, mrpp_cat$sev)

lichen.mrpp$Pvalue

summary(lichen.mrpp)

# Save and change plotting parameters
def.par <- par(no.readonly = TRUE)
layout(matrix(1:2,nr=1))

plot(lichen.ord <- metaMDS(mrpp_data), type="text", display="sites" )
ordihull(lichen.ord, mrpp_cat$sev, label = T)

with(lichen.mrpp, {
  fig.dist <- hist(boot.deltas, xlim=range(c(delta,boot.deltas)), 
                 main="Test of Differences Among Groups")
  abline(v=delta); 
  text(delta, 2*mean(fig.dist$counts), adj = -0.5,
     expression(bold(delta)), cex=1.5 )  }
)
par(def.par)

## meandist 
# calculates a matrix of mean within-cluster dissimilarities (diagonal) and between-cluster dissimilarities (off-diagonal elements)
lichen.md <- meandist(vegdist(mrpp_data), mrpp_cat$sev)
lichen.md
summary(lichen.md)
plot(lichen.md)

library(asbio)
Mrpp(mrpp_data, mrpp_cat$sev)
??Mrpp
library(Mrpp)

library(vegan)
data(dune)
data(dune.env)
asbio::Mrpp(dune, dune.env$Management)

help(package = asbio)
# can I install package asbio v. 0.3-1? Or check for later versions?





d <- dist(as.matrix(plant_mat_pa_w))   # find distance matrix 
hc <- hclust(d)                # apply hirarchical clustering 
plot(hc)  


hc <- agnes(d)
pltree(hc, cex = 0.6, hang = -1, main = "Dendrogram of agnes")

hc <- diana(d)
pltree(hc, cex = 0.6, hang = -1, main = "Dendrogram of agnes")

library(cluster)
install.packages("factoextra", "dendextend")
library(factoextra) # clustering visualization
library(dendextend)


df <- scale(all_spp_w_pa %>% remove_rownames %>% column_to_rownames(var="site_id") %>% dplyr::select(-c(CHLA, ABPR, ALRH, SALU)))

df <- scale(bird_mat_w %>% remove_rownames %>% column_to_rownames(var="site_id")) #%>% select(-c(CHLA, ABPR, ALRH, SALU)))

# Dissimilarity matrix
d <- dist(df, method = "euclidean")

# Hierarchical clustering using Complete Linkage
hc1 <- hclust(d, method = "complete" )

# Plot the obtained dendrogram
plot(hc1, cex = 0.6, hang = -1)

# Compute with agnes
hc2 <- agnes(df, method = "complete")

# Agglomerative coefficient
hc2$ac

# methods to assess
m <- c( "average", "single", "complete", "ward")
names(m) <- c( "average", "single", "complete", "ward")

# function to compute coefficient
ac <- function(x) {
  agnes(df, method = x)$ac
}

map_dbl(m, ac)

hc3 <- agnes(df, method = "ward")
pltree(hc3, cex = 0.6, hang = -1, main = "Dendrogram of agnes")

# compute divisive hierarchical clustering
hc4 <- diana(df)

# Divise coefficient; amount of clustering structure found
hc4$dc

# plot dendrogram
pltree(hc4, cex = 0.6, hang = -1, main = "Dendrogram of diana")

# Ward's method
hc5 <- hclust(d, method = "ward.D2" )

# Cut tree into 4 groups
sub_grp <- cutree(hc5, k = 3)

# Number of members in each cluster
table(sub_grp)

dfff<-bird_dat_w %>%
  mutate(cluster = sub_grp) %>%
  select(c(1, sev, tree_cov, tsf_2cat:sev_cov, cluster))

plot(hc5, cex = 0.6)
rect.hclust(hc5, k = 9, border = 2:5)

fviz_cluster(list(data = df, cluster = sub_grp))

```


### PERMANOVA
#### Plants
A permanova analysis shows that there is a strong effect of severity on plants (only native species used in the analysis), and a weaker effect of TSF. A pairwise comparison indicates that there is no statistically significant difference between unburnt and LS sites, but that HS sites differ in species composition. Further analysis incorporating TSF is needed.  

```{r echo = F}
# Main PERMANOVA
#comp.data$tsf <- as.numeric(comp.data$tsf)
plant_matrix$site_id <- as.character(plant_matrix$site_id)

#plant_matrix_sub <- plant_matrix %>% filter(sev == "")
#site_data_sub <- site_data %>% filter(sev == "h")

#comp.data <- left_join(plant_matrix, site_data, by = "site_id") %>% 
#  filter(sev == "h")# %>% 
#  mutate(sev2 = ifelse(sev == 'h' , "h", "lu"))

#comp.data$sev <- as.character(comp.data$sev)
#comp.data$tsf <- as.numeric(comp.data$tsf)

# Pick among the following
## Is there an effect of severity and time since fire?
## log transforming like NMDS
## Using all tsf_cat and sev, we see a strong effect of sev, and a marginal effect of tsf_cat

comp.data <- left_join(plant_matrix, site_data, by = "site_id")

comp.data <- plant_dat_cov_w
comp.sub <- subset(comp.data, select = ACAM:WHMO)
comp.env <- subset(comp.data, select = c(sev:tsf_cat))
attach(comp.env)

kable(adonis2(log1p(comp.sub) ~ sev * tsf_cat, permutations = 9999), caption = "Table 2: PERMANOVA results", digits = 2) %>%
  kable_styling()

## Using the filter function, can look at only l of h sev. There is no effect of tsf_cat in l, but there is an effect in h.
#comp.data <- left_join(plant_matrix, site_data, by = "site_id") %>% 
#  filter(sev == "h")
#comp.sub <- subset(comp.data, select = ACTR:WHMO)
#comp.env <- subset(comp.data, select = c(fire:tsf_cat))
#attach(comp.env)
#adonis2(comp.sub ~ sev * tsf_cat)


##the second term is because i'm trying to compare LS and UN at younger stages. But this code excludes UN so it doesn't work.
#comp.sub <- subset(comp.data, (tsf < 13 | sev != "h"), select =ACTR:WHMO) # might need to use dplyr::select; also replace ACTR:WHMO with ABCO:UMCA if using trees or trees and shrubs)
#comp.env <- subset(comp.data, (tsf < 13 | sev != "h"), select = c(fire:tsf_cat))
#attach(comp.env)
#adonis2(comp.sub ~ sev * tsf_cat)

```


```{r eval = F, include = F}
## Pairwise PERMANOVA function (pt.2)
### below is an attempt to show Df, but output is screwy

pairwise.adonis <- function(x, factors, sim.method = "bray", p.adjust.m ="bonferroni") {
  library(vegan)
  co <- combn(unique(factors), 2)
  pairs <- c()
  Df <- c()
  F.Model <- c()
  R2 <- c()
  p.value <- c()

  for (elem in 1:ncol(co)) {
    ad <- adonis(x[factors %in% c(co[1, elem], co[2, elem]), ] ~ factors[factors %in% c(co[1, elem], co[2, elem])], method = sim.method)
    pairs <- c(pairs, paste(co[1, elem], "vs", co[2, elem]))
    Df <- c(Df, ad$aov.tab[1, ])
    F.Model <- c(F.Model, ad$aov.tab[1, 4])
    R2 <- c(R2, ad$aov.tab[1, 5])
    p.value <- c(p.value, ad$aov.tab[1, 6])
  }
  p.adjusted <- p.adjust(p.value, method = p.adjust.m)
  pairw.res <- data.frame(pairs, Df, F.Model, R2, p.value, p.adjusted)
  return(pairw.res)
}
```

```{r eval = F, include = F}
## Pairwise PERMANOVA function (run)
# 1 = High severity; 2 = Low severity; 3 = Unburnt.
comp.data <- comp.data %>% 
    mutate(sev2 = ifelse(sev == 'h' , "1", ifelse(sev == 'l' , "2", "3")))

pair <- pairwise.adonis(comp.sub, comp.data$sev2) %>% 
  dplyr::select(pairs, F.Model, p.adjusted)

kable(pair, caption = "Table 3: pairwise PERMANOVA results", digits = 4) %>%
  kable_styling()

```

```{r eval=F, echo = F, include = F}
# Pairwise comparison for herbaceous plants using TSF categories.
# here I'm setting the TSF categories; change as necessary.
site_data <- site_data_read %>% 
    mutate(tsf_cat = ifelse(is.na(tsf), "4", "3"))
site_data$tsf_cat <-as.numeric(as.character(site_data$tsf_cat))
site_data$tsf_cat[site_data$tsf < 15] <-2
site_data$tsf_cat[site_data$tsf < 10] <-1

plant_matrix$site_id <- as.character(plant_matrix$site_id)
site_data$site_id <- as.character(site_data$site_id)

plant_dat <- left_join(plant_data, plant_names, by = "species") %>%
  filter(native_status == "native" & (form == "herb" | form == "grass")) %>%
  dplyr::select(site_id, species, cover)
plant_matrix <- spread(data = plant_dat, key = species, value = cover, fill = 0)

comp.data <- left_join(plant_matrix, site_data, by = "site_id") %>% 
  mutate(sev2 = ifelse(sev == 'h' , "1", ifelse(sev == 'l' , "2", "3"))) #%>%  %>% filter(sev == "h")
comp.sub <- subset(comp.data, select = ACAM:YAMI) #ADBI:WOFI)

kable(pairwise.adonis(comp.sub, comp.data$sev2))

```

#### Insects
Again, the PERMANOVA shows a significant impact of severity and TSF on insect communities (there is a barely significant interaction). Pairwise analysis reveals no detectable difference between UN and LS but HS is significantly different. Using the tsf and sev together to distinguish between recent HS and older HS versus LS/UN, there is no detectable difference between older HS and LS/UN, but there is also not detectable difference between recent and older HS.
```{r echo = F}
# Main PERMANOVA
insect_matrix$site_id <- as.character(insect_matrix$site_id)

comp.data <- insect_mrpp_d
comp.sub <- subset(comp.data, select = 3:19)
comp.env <- subset(comp.data, select = c(fire:tsf_cat))
attach(comp.env)

kable(adonis2(log1p(comp.sub) ~ sev * tsf_cat, permutations = 9999), caption = "Table 2: PERMANOVA results", digits = 2) %>%
  kable_styling()
```

```{r include=F}
### Pairwise PERMANOVA (code from https://www.researchgate.net/post/How_can_I_do_PerMANOVA_pairwise_contrasts_in_R)
pairwise.adonis <- function(x, factors, sim.method = "bray", p.adjust.m ="bonferroni") {
  library(vegan)
  co <- combn(unique(factors), 2)
  pairs <- c()
  F.Model <- c()
  R2 <- c()
  p.value <- c()

  for (elem in 1:ncol(co)) {
    ad <- adonis(x[factors %in% c(co[1, elem], co[2, elem]), ] ~ factors[factors %in% c(co[1, elem], co[2, elem])], method = sim.method)
    pairs <- c(pairs, paste(co[1, elem], "vs", co[2, elem]))
    F.Model <- c(F.Model, ad$aov.tab[1, 4])
    R2 <- c(R2, ad$aov.tab[1, 5])
    p.value <- c(p.value, ad$aov.tab[1, 6])
  }
  p.adjusted <- p.adjust(p.value, method = p.adjust.m)
  pairw.res <- data.frame(pairs, F.Model, R2, p.value, p.adjusted)
  return(pairw.res)
}

```


```{r eval = F, include = F}
## Pairwise PERMANOVA function (run)
# 1 = High severity; 2 = Low severity; 3 = Unburnt.
comp.data <- comp.data %>% 
    mutate(sev2 = ifelse(sev == 'h' , "1", ifelse(sev == 'l' , "2", "3")))

pair <- pairwise.adonis(comp.sub, comp.data$sev2) %>% 
  dplyr::select(pairs, F.Model, p.adjusted)

kable(pair, caption = "Table 4: pairwise PERMANOVA results for insects", digits = 2) %>%
  kable_styling()

# this time using tsf_cat
#comp.data <- comp.data %>% 
#    mutate(tsf_cat2 = ifelse(sev_tsf == 'h-2'|sev_tsf == "h-1", "1", ifelse(sev_tsf == 'h-3' , "2", "3")))

comp.data <- comp.data %>% 
    mutate(tsf_cat2 = ifelse(sev_tsf == 'h-1', "1", ifelse(sev_tsf == 'h-2' , "2", ifelse(sev_tsf == "h-3", "3", "4"))))

pair <- pairwise.adonis(comp.sub, comp.data$tsf_cat2) %>% 
  dplyr::select(pairs, F.Model, p.adjusted)

kable(pair, caption = "Table 4: pairwise PERMANOVA results for insects", digits = 2) %>%
  kable_styling()

```

```{r eval=F, echo = F, include = F}
# Pairwise comparison for herbaceous insects using TSF categories.
# here I'm setting the TSF categories; change as necessary.
site_data <- site_data_read %>% 
    mutate(tsf_cat = ifelse(is.na(tsf), "4", "3"))
site_data$tsf_cat <-as.numeric(as.character(site_data$tsf_cat))
site_data$tsf_cat[site_data$tsf < 15] <-2
site_data$tsf_cat[site_data$tsf < 10] <-1

insect_matrix$site_id <- as.character(insect_matrix$site_id)
site_data$site_id <- as.character(site_data$site_id)

insect_dat <- left_join(insect_data, insect_names, by = "species") %>%
  filter(native_status == "native" & (form == "herb" | form == "grass")) %>%
  dplyr::select(site_id, species, cover)
insect_matrix <- spread(data = insect_mat, key = species, value = cover, fill = 0)

comp.data <- left_join(insect_matrix, site_data, by = "site_id") %>% 
  mutate(sev2 = ifelse(sev == 'h' , "1", ifelse(sev == 'l' , "2", "3"))) #%>%  %>% filter(sev == "h")
comp.sub <- subset(comp.data, select = ADBI:WOFI)

kable(pairwise.adonis(comp.sub, comp.data$sev2))

```

### Indicator species analysis

#### Plants
The indicator species analysis reveals which species are strongly tied to HS and LS/UN habitats (the latter are grouped since the PERMANOVA suggested they were not very different). Further exploration of species preferences using linear models is in progress.  

In brief, species associated with high severity burns include native grasses, shrubs, and annuals, whereas species of closed-canopy stands include sword fern, vanilla leaf, and pathfinder. Unsurprisingly, high cover of Douglas-fir is indicative of low severity fire.  

```{r echo = F}
### Community data matrix
plant_data_filtered <- left_join(plant_data, site_data, by = "site_id") 

plant_dat <- left_join(plant_data_filtered, plant_names, by = "species") %>%
  filter(native_status == "native") %>%
  mutate(sev_tsf = paste(sev, tsf_cat, sep = "")) %>% 
  mutate(sev2 = ifelse(sev == 'h' , "h", "lu")) 

plant_dat_select  <- plant_dat %>%
  dplyr::select(site_id, species, cover, sev, sev_tsf, sev2)

plant_matrix1 <- spread(data = plant_dat_select, key = species, value = cover, fill = 0)

# getting rid of species detected in 0 plots (for some reason there appears to be no cover recorded for ALRH and CESA, but they were both found at only one plot so no big deal)
plant_matrix1 <- plant_matrix1 %>% 
  bind_rows(summarise_all(., funs(if(is.numeric(.)) sum(.) else "Total")))

plant_matrix1 <- plant_matrix1[, (plant_matrix1[49, ]) > 5]
plant_matrix1 <- plant_matrix1[-49, ]


### Vector for site classification
### severity

bla <- plant_mat_cov_w %>% 
  bind_rows(summarise_all(., funs(if(is.numeric(.)) sum(.) else "Total")))

bla <- bla[, (bla[111, ]) > 0]
bla <- bla[-111, ]



data <- bla[,c(2:178)]
groups1 <- as.vector(plant_dat_pa_w$sev)

#groups1 = as.vector(plant_dat$tsf_cat2)
#groups1 = as.vector(plant_dat$sev_tsf)

Indval_out1 <- indval(data, groups1, numitr=10000)
gr <- Indval_out1$maxcls[Indval_out1$pval<=0.05]
iv <- Indval_out1$indcls[Indval_out1$pval<=0.05]
pv <- Indval_out1$pval[Indval_out1$pval<=0.05]
fr <- apply(data>0, 2, sum)[Indval_out1$pval<=0.05]
indvalsummary <- data.frame(group=gr, indval=iv, pvalue=pv, freq=fr)
indvalsummary1 <- indvalsummary[order(indvalsummary$group, -indvalsummary$indval),]
prob.corrected1 = p.adjust(Indval_out1$pval, "BH") # correct p.value for multiple testing
#kable(indvalsummary1)

indvalsummary1 <- cbind(rownames(indvalsummary1), data.frame(indvalsummary1, row.names=NULL))

names(indvalsummary1)[1]<-"species"

indval <- left_join(indvalsummary1, plant_names, by = "species") %>% 
  dplyr::select(full_name, group, indval, pvalue) %>% 
  mutate("Indicator value" = indval * 100) %>% 
  filter(indval >= 0.25) %>% 
  filter(pvalue < 0.05)

indval_hs <- indval %>% 
  filter(group =="1") %>% 
  dplyr::select(full_name, "Indicator value")

indval_ls <- indval %>% 
  filter(group =="2")%>% 
  dplyr::select(full_name, "Indicator value")

indval_3 <- indval %>% 
  filter(group =="3")%>% 
  dplyr::select(full_name, "Indicator value")

names(indval_hs)[1]<-"Indicators of HS"
names(indval_ls)[1]<-"Indicators of LS/UN"
names(indval_3)[1]<-"Indicators of 3"

#names(spp_diff2)[1]<-"Species found only in LS/UN"
indval_table <- list(indval_hs, indval_ls, indval_3)

kable(indval_table, caption = "Table 3: Indicator species for different habitats", digits = 0) %>%
  kable_styling()

#write.csv(indvalsummary1, file = "indvalsummary1.csv")
#write.csv(prob.corrected1, file = "prob.corrected1.csv")
```


#### Birds 
```{r echo = F}
bird_matrix1 <- bird_matrix1 %>% 
  mutate(sev2 = ifelse(sev == 'h' , "h", "lu"))

### Vector for site classification
### severity
data <- bird_matrix1[,c(28:83)]
groups1 <- as.vector(bird_matrix1$sev2)

#groups1 = as.vector(plant_dat$tsf_cat2)
#groups1 = as.vector(plant_dat$sev_tsf)

Indval_out1 <- indval(data, groups1, numitr=10000)
gr <- Indval_out1$maxcls[Indval_out1$pval<=0.05]
iv <- Indval_out1$indcls[Indval_out1$pval<=0.05]
pv <- Indval_out1$pval[Indval_out1$pval<=0.05]
fr <- apply(data>0, 2, sum)[Indval_out1$pval<=0.05]
indvalsummary <- data.frame(group=gr, indval=iv, pvalue=pv, freq=fr)
indvalsummary1 <- indvalsummary[order(indvalsummary$group, -indvalsummary$indval),]
prob.corrected1 = p.adjust(Indval_out1$pval, "BH") # correct p.value for multiple testing
#kable(indvalsummary1)

indvalsummary1 <- cbind(rownames(indvalsummary1), data.frame(indvalsummary1, row.names=NULL))

names(indvalsummary1)[1]<-"species"

indval <- indvalsummary1 %>% 
  dplyr::select(species, group, indval, pvalue) %>% 
  mutate("Indicator value" = indval * 100) %>% 
  filter(indval > 0.245) %>% 
  filter(pvalue < 0.05)

indval_hs <- indval %>% 
  filter(group =="1") %>% 
  dplyr::select(species, "Indicator value")

indval_ls <- indval %>% 
  filter(group =="2")%>% 
  dplyr::select(species, "Indicator value")

names(indval_hs)[1]<-"Indicators of HS"
names(indval_ls)[1]<-"Indicators of LS/UN"

#names(spp_diff2)[1]<-"taxon found only in LS/UN"
indval_table <- list(indval_hs, indval_ls)

kable(indval_table, caption = "Table 3: Bird indicator species for different habitats", digits = 0) %>%
  kable_styling()
#write.csv(indvalsummary1, file = "indvalsummary1.csv")
#write.csv(prob.corrected1, file = "prob.corrected1.csv")
```


#### Insects

There are no species that are indicators of LS/UN. Several orders/suborders have a high indicator value for HS though: Hemiptera (Homoptera specifically), Coleoptera, and Hymenoptera (Aculeata specifically). Brachyceran flies are also indicators of HS (but not Diptera as a whole).
```{r echo = F}
# Putting Hemiptera and Hymenoptera back together, the indicator value stays the same as when splitting, but when split, Homoptera and Aculeata replace the order (respectively). Splitting Diptera takes it out of the final output altogether.

### Community data matrix
insect_dat_ind1 <- insect_dat_l %>%  
  mutate(sev_tsf = paste(sev, tsf_cat, sep = "")) %>% 
  mutate(sev2 = ifelse(sev == 'h' , "h", "lu")) %>% 
  mutate(taxon = recode(taxon, Aculeata="Hymenoptera")) %>% 
  mutate(taxon = recode(taxon, Parasitica="Hymenoptera")) %>% 
  mutate(taxon = recode(taxon, Homoptera ="Hemiptera")) %>% 
  mutate(taxon = recode(taxon, Heteroptera ="Hemiptera"))

insect_dat_ind <- as.data.table(insect_dat_ind1)[, sum(number), by = .(site_id, taxon)] %>% 
  left_join(., site_data) %>%
  mutate(sev_tsf = paste(sev, tsf_cat, sep = "")) %>% 
  mutate(sev2 = ifelse(sev == 'h' , "h", "lu")) 

#insect_dat_select  <- insect_dat_ind %>%
#  dplyr::select(site_id, taxon, V1, sev, sev_tsf, sev2)

insect_matrix1 <- spread(data = insect_dat_ind, key = taxon, value = V1, fill = 0)

# getting rid of taxon detected in few plots
#insect_matrix1 <- insect_matrix1 %>% 
#  bind_rows(summarise_all(., funs(if(is.numeric(.)) sum(.) else "Total")))

#insect_matrix1 <- insect_matrix1[, (insect_matrix1[49, ]) > 5]
#insect_matrix1 <- insect_matrix1[-49, ]


### Vector for site classification
### severity
data <- insect_matrix1[,-c(1:29)]
groups1 <- as.vector(insect_matrix1$sev2)

#groups1 = as.vector(insect_dat$tsf_cat2)
#groups1 = as.vector(insect_dat$sev_tsf)

Indval_out1 <- indval(data, groups1, numitr=10000)
gr <- Indval_out1$maxcls[Indval_out1$pval<=0.05]
iv <- Indval_out1$indcls[Indval_out1$pval<=0.05]
pv <- Indval_out1$pval[Indval_out1$pval<=0.05]
fr <- apply(data>0, 2, sum)[Indval_out1$pval<=0.05]
indvalsummary <- data.frame(group=gr, indval=iv, pvalue=pv, freq=fr)
indvalsummary1 <- indvalsummary[order(indvalsummary$group, -indvalsummary$indval),]
prob.corrected1 = p.adjust(Indval_out1$pval, "BH") # correct p.value for multiple testing
#kable(indvalsummary1)

indvalsummary1 <- cbind(rownames(indvalsummary1), data.frame(indvalsummary1, row.names=NULL))

names(indvalsummary1)[1]<-"taxon"

indval <- indvalsummary1 %>% 
  dplyr::select(taxon, group, indval, pvalue) %>% 
  mutate("Indicator value" = indval * 100) %>% 
  filter(indval > 0.245) %>% 
  filter(pvalue < 0.05)

indval_hs <- indval %>% 
  filter(group =="1") %>% 
  dplyr::select(taxon, "Indicator value")

indval_ls <- indval %>% 
  filter(group =="2")%>% 
  dplyr::select(taxon, "Indicator value")

names(indval_hs)[1]<-"Indicators of HS"
names(indval_ls)[1]<-"Indicators of LS/UN"

#names(spp_diff2)[1]<-"taxon found only in LS/UN"
indval_table <- list(indval_hs, indval_ls)

kable(indval_table, caption = "Table 3: Indicator taxon for different habitats", digits = 0) %>%
  kable_styling()

#write.csv(indvalsummary1, file = "indvalsummary1.csv")
#write.csv(prob.corrected1, file = "prob.corrected1.csv")
```

### Growth-stage optimization
#### GSO: plants

In a nutshell, the growth-stage optimization method (GSO) uses an algorithm to produce optimal proportions of different habitats that would maximize the abundance of all species across a landscape. It has been used in Australia to produce actionable targets for fire managers to achieve the best results for biodiversity. In this case, I divided the landscape into high severity (HS), low severity (LS), and unburnt (UN), with two age categories for HS and LS: 4-12 years since fire (1), and 16-31 years since fire (2). The GSO for plants suggests that ~80% of the landscape should be burnt at high severity, with the rest divided into low severity and unburnt (Fig. 3), but see discussion for how to interpret this startling result.

```{r Read in plant data, echo=FALSE, include=F}
## Packages needed throughout
#library (plyr)
#library (dplyr)
#library (reshape2)

#setwd("C:/Users/Documents/Example")

# w = data are in wide format; a, b, c, d keep growth stages in order

#select just the species column
plant_spp <- read.csv("data/plant_data_natives.csv", header=TRUE) %>% 
  dplyr::select(species) 

#now make a list of all species
plant_dat <- read.csv("data/plant_data_natives.csv", header=TRUE)

plant_dat_UN <- plant_dat %>% 
  filter(sev == "u") %>% 
  dplyr::select(species)

complete_spp <- union(plant_spp, plant_dat_UN)

## Unburnt data: filter to get only UN; keep only species and cover columns
plant_dat_UN <- plant_dat %>% 
  filter(sev == "u") %>% 
  dplyr::select(species, cover, site_id)  %>% 
  arrange(species)

#plant_dat_UN <- full_join(plant_dat_UNa, plant_dat_UNb, by = "species")%>% 
#  arrange(species)

# make it wide format
plant_dat_UN <- reshape(plant_dat_UN, idvar = "species", timevar = "site_id", direction = "wide")

# add species not detected to get same length vectors
#plant_dat_UN1 

ar.w <- full_join(complete_spp, plant_dat_UN, by = "species") %>% 
  replace(., is.na(.), "0")

## LS data (less than 15 yrs): filter to get only LS, TSF less than 14 years, keep only species and cover columns
plant_dat_LS1 <- plant_dat %>% 
  filter(sev == "l")

plant_dat_LS1$fire_yr <- as.character(plant_dat_LS1$fire_yr)
plant_dat_LS1$fire_yr <- as.numeric(plant_dat_LS1$fire_yr)
plant_dat_LS1$tsf <- 2018-plant_dat_LS1$fire_yr

plant_dat_LS1 <- plant_dat_LS1 %>% 
  filter(tsf < 15) %>% 
  dplyr::select(species, cover, site_id) %>% 
  arrange(species)
  
plant_dat_LS1 <- reshape(plant_dat_LS1, idvar = "species", timevar = "site_id", direction = "wide")

# add species not detected to get same length vectors
#plant_dat_LS1 

be.w <- full_join(complete_spp, plant_dat_LS1, by = "species") %>% 
  replace(., is.na(.), "0")

## LS data (less than 10 yrs; 11 sites total): filter to get only LS, TSF more than 10 years, keep only species and cover columns
plant_dat_LS2 <- plant_dat %>% 
  filter(sev == "l")

plant_dat_LS2$fire_yr <- as.character(plant_dat_LS2$fire_yr)
plant_dat_LS2$fire_yr <- as.numeric(plant_dat_LS2$fire_yr)
plant_dat_LS2$tsf <- 2018-plant_dat_LS2$fire_yr

plant_dat_LS2 <- plant_dat_LS2 %>% 
  filter(tsf > 15) %>% 
  dplyr::select(species, cover, site_id) %>% 
  arrange(species)
  
plant_dat_LS2 <- reshape(plant_dat_LS2, idvar = "species", timevar = "site_id", direction = "wide")

# add species not detected to get same length vectors
#plant_dat_LS1 

xx.w <- full_join(complete_spp, plant_dat_LS2, by = "species") %>% 
  replace(., is.na(.), "0")

# HS data (less than 14 yrs; ?? sites): filter to get only LS, TSF less than 14 years, keep only species and cover columns
plant_dat_HS1 <- plant_dat %>% 
  filter(sev == "h")

plant_dat_HS1$fire_yr <- as.character(plant_dat_HS1$fire_yr)
plant_dat_HS1$fire_yr <- as.numeric(plant_dat_HS1$fire_yr)
plant_dat_HS1$tsf <- 2018-plant_dat_HS1$fire_yr

plant_dat_HS1 <- plant_dat_HS1 %>% 
  filter(tsf < 15) %>% 
  dplyr::select(species, cover, site_id) %>% 
  arrange(species)

plant_dat_HS1 <- reshape(plant_dat_HS1, idvar = "species", timevar = "site_id", direction = "wide")

#plant_dat_LS11 

cm.w <- full_join(complete_spp, plant_dat_HS1, by = "species") %>% 
  replace(., is.na(.), "0")

# HS data (more than 10 yrs): filter to get only LS, TSF more than 14 years, keep only species and cover columns
plant_dat_HS2 <- plant_dat %>% 
  filter(sev == "h")

plant_dat_HS2$fire_yr <- as.character(plant_dat_HS2$fire_yr)
plant_dat_HS2$fire_yr <- as.numeric(plant_dat_HS2$fire_yr)
plant_dat_HS2$tsf <- 2018-plant_dat_HS2$fire_yr

plant_dat_HS2 <- plant_dat_HS2 %>% 
  filter(tsf > 15) %>% 
  dplyr::select(species, cover, site_id) %>% 
  arrange(species)

plant_dat_HS2 <- reshape(plant_dat_HS2, idvar = "species", timevar = "site_id", direction = "wide")

#plant_dat_HS22 

dl.w <- full_join(complete_spp, plant_dat_HS2, by = "species") %>% 
  replace(., is.na(.), "0")
```

```{r Melt plant data, warning=FALSE, results="hide", echo =F, include = F}
#Convert the data format from wide to long (such that there are only three columns: species, site, value), and order by species. Then delete the site column. 

melt.sort <- function (x){
  m <- melt(x, id.vars="species") # convert to long
  s <- m[order(m$species),] # order by species
  s$variable <- NULL # delete site column
  return(s)}

# l = data are in long format
ar.l <- melt.sort(ar.w) %>% 
  mutate(value = as.numeric(value))
be.l <- melt.sort(be.w) %>% 
  mutate(value = as.numeric(value))
xx.l <- melt.sort(xx.w) %>% 
  mutate(value = as.numeric(value))
cm.l <- melt.sort(cm.w) %>% 
  mutate(value = as.numeric(value))
dl.l <- melt.sort(dl.w) %>% 
  mutate(value = as.numeric(value))

#Now data are ready for resampling.  Use a different method for abundance data (resampling with replacement) and presence-absence data (random number generation from the binomial distribution).  Both methods produce a matrix named `zdata`.  

#An extra step is required for data containing subsamples -- see below.

```


```{r Plant Abundance, echo=FALSE, include = F}
#### Abundance data
#Use this method for abundance datasets which don't contain subsamples.

res.ab <- function (z){(ddply (z, .(species), function(x) { 
  rs1 <- sample(x$value, replace = TRUE) # resample raw data
  bm <- mean(rs1) # big mean is the mean of the resampled means  
  data.frame(bm=bm)}
  ))
}

# Run 'res.ab' on each growth stage:
ar.r <- rdply (1000, res.ab (ar.l))
be.r <- rdply (1000, res.ab (be.l))
xx.r <- rdply (1000, res.ab (xx.l))
cm.r <- rdply (1000, res.ab (cm.l))
dl.r <- rdply (1000, res.ab (dl.l))

# Combine to a matrix:
zdata <- cbind (ar.r$.n, ar.r$bm, be.r$bm, xx.r$bm, cm.r$bm, dl.r$bm) # ar.r$.n = row labels
```

```{r Combine plant data, echo = FALSE,include =F}
### Convert the matrix to a data frame and save it.
#The same method applies to all data types from this point.

mdata <- apply (zdata, 2, function (x) ifelse (x>0, x, 0.001)) 
# replace zeros for geometric mean calculation
rdata <- data.frame(mdata) # convert to a data frame
names(rdata) <- c("group", "AR", "BE", "XX", "CM", "DL")
rdata$group <- factor(rdata$group) # rdata is ready for the GSO function
```


```{r GSO function for plants, echo = FALSE, warning=FALSE, message = F}
## 2 Run the GSO function

#The nloptr function maximises the geometric mean by changing the growth stage proportions given bounds constraints (parameters (GS proportions) must be between 0 and 1), and an inequality constraint -- the parameters must sum to 1. 

library(nloptr)
library(plyr)
library(dplyr)
## Handy helper function, constructing a n-length vector c(k, ..., k)
const.vector <- function(n, k) {
  return( sapply(seq_len(n), function (x) return(k)) )
}
## The original objective (assumes columns [AR, BE, CM, DL] appear in that order) 
geomean.fun <- function(x, species) {
  logsum <- sum(log(apply(species, 1, function (row) { sum(x*row) })))
  exp(logsum/(dim(species)[1]))
}
## The reformulated objective (nloptr expects a minimization problem)
geomean.obj <- function (x, species) {
  logsum <- sum(log(apply(species, 1, function (row) { sum(x*row) })))
  return(-logsum)
}
## The gradient of geomean.obj.
## Returns a vector containing the partial derivatives wrt each var.
geomean.grad <- function(x, species) {
  contrib <- apply(species, 1, function(row) { -row/(sum(x*row)) })
  grad <- apply(contrib, 1, sum)  # sum the contribution vectors to get the gradient
  return(grad)
}
## Inequality constraints.
eval.cs <- function(x, species) {
  return( c(
    sum(x) - 1,  # Growth stage proportions don't exceed 1
    0.0 - x[1]   # Recently burnt proportion is at least 0.0 (i.e. AR >= 0.0)
  ))
}
eval.cs.jac <- function(x, species) { # Jacobian (partial derivative matrix) for inequalities
  return( rbind(
    const.vector(length(x), 1),
    c(-1, const.vector(length(x)-1, 0))
    ))
}
## Pass all to nloptr 
gso <- function(species){
  require(nloptr)
  require(dplyr)
  x0 <- const.vector(5, 1/5)  # edit according to the number of growth stages
  run <- nloptr(x0 = x0, species = species, eval_f=geomean.obj, eval_grad_f = geomean.grad,
                lb = const.vector(length(x0), 0),
                ub = const.vector(length(x0), 1),
                eval_g_ineq = eval.cs,
                eval_jac_g_ineq = eval.cs.jac,
                opts = list("algorithm" = "NLOPT_LD_MMA",
                            "maxeval"=1000,
                            "ftol_rel"=1.0e-15,
                            "xtol_rel"=1.0e-8,
                            "print_level"=0
                            ))
  res <- rbind(c(run$solution, geomean.fun(run$solution, species)))
  colnames(res) <- c("AR", "BE", "XX", "CM", "DL", "geom") # edit according to growth stages
  res <- data.frame(res)
  return (res)
}

# Apply to the 1000 resampled communities (defined by group)
out <- ddply(rdata[2:6], .(rdata$group), gso) # takes a few minutes to run

names(out)[1]<-"num"
names(out)[2]<-"UN"
names(out)[3]<-"LS1"
names(out)[4]<-"LS2"
names(out)[5]<-"HS1"
names(out)[6]<-"HS2"

UN <- mean(out$UN) *100
LS1 <- mean(out$LS1) *100
LS2 <- mean(out$LS2) *100
HS1 <- mean(out$HS1) *100
HS2 <- mean(out$HS2) *100

GSO_tbl <- list(UN, LS1, LS2, HS1, HS2) 

names(GSO_tbl)[1]<-"UN"
names(GSO_tbl)[2]<-"LS1"
names(GSO_tbl)[3]<-"LS2"
names(GSO_tbl)[4]<-"HS1"
names(GSO_tbl)[5]<-"HS2"

GSO_tbl <- data.frame(GSO_tbl)

#kable(GSO_tbl, caption = "Table 5: GSO results (plants)", digits = 0, col.names = c("UN", 
#                                                                           "LS1", 
#                                                                           "LS2", 
#                                                                           "HS1", 
#                                                                           "HS2")) %>%
#  kable_styling()

out_long <- gather(out, severity) %>% 
  filter(severity != "num" & severity != "geom") 

out_long$value <- as.numeric(out_long$value)

out_long <- out_long %>% 
  mutate("prop" = value*100)

# summarize data
data_summary <- function(data, varname, groupnames){
  require(plyr)
  summary_func <- function(x, col){
    c(mean = mean(x[[col]], na.rm=TRUE),
      sd = sd(x[[col]], na.rm=TRUE))
  }
  data_sum<-ddply(data, groupnames, .fun=summary_func,
                  varname)
  data_sum <- rename(data_sum, c("mean" = varname))
 return(data_sum)
}

df2 <- data_summary(out_long, varname="prop", groupnames="severity")

#plot
hist <- ggplot(df2, aes(x = severity, y = prop)) +
  geom_histogram(stat = "identity") +
  labs(title = "Figure 3: Optimal habitat proportions (plants)", y = "% of landscape", x = "habitat") +
  geom_errorbar(aes(ymin=prop-sd, ymax=prop+ sd), width=.2) 
hist


# Or apply to a single community (for example, the raw data)
#out <- gso (rdata)

#write.table (out, "gso_result.txt", col.names=TRUE, row.names=FALSE)
```

#### GSO: birds

The GSO for birds is similar to that for plants (Fig. 4). The optimal proportion of habitats to maximize biodiversity would include a large proportion of HS fire, and closed canopy (LS or UN) representing roughly a quarter of the landscape.

```{r Bird Read in data, echo=FALSE, include = F}
## Packages needed throughout
#library (plyr)
#library (dplyr)
#library (reshape2)

# create list of all species; NOTE this is better code than for plants above, might consider changing that
bird_spp <- bird_dat_long %>% 
  dplyr::select(species, pa)

bird_spp <-  unique(bird_spp[ , c(1,2)]) %>% 
  dplyr::select(species)

# Add site data to detection data
bird_dat_long$site_id <- as.character(bird_dat_long$site_id)
bird_dat_long <- left_join(bird_dat_long, site_data, by = "site_id")

## Unburnt data: filter to get only UN; keep only species and cover columns
bird_dat_UN <- bird_dat_long %>% 
  filter(sev == "u") %>% 
  dplyr::select(species, pa, site_id)  %>% 
  arrange(species)

# make it wide format
bird_dat_UN <- reshape(bird_dat_UN, idvar = "species", timevar = "site_id", direction = "wide")

# add species not detected to get same length vectors
ar.w <- full_join(bird_spp, bird_dat_UN, by = "species") %>% 
  replace(., is.na(.), "0")

## LS data (less than 11 yrs): filter to get only LS, TSF less than 14 years, keep only species and cover columns
bird_dat_LS1 <- bird_dat_long %>% 
  filter(sev == "l")

bird_dat_LS1$fire_yr <- as.character(bird_dat_LS1$fire_yr)
bird_dat_LS1$fire_yr <- as.numeric(bird_dat_LS1$fire_yr)
bird_dat_LS1$tsf <- 2018-bird_dat_LS1$fire_yr

bird_dat_LS1 <- bird_dat_LS1 %>% 
  filter(tsf < 15) %>% 
  dplyr::select(species, pa, site_id) %>% 
  arrange(species)
  
bird_dat_LS1 <- reshape(bird_dat_LS1, idvar = "species", timevar = "site_id", direction = "wide")

# add species not detected to get same length vectors
be.w <- full_join(bird_spp, bird_dat_LS1, by = "species") %>% 
  replace(., is.na(.), "0")

## LS data (less than 10 yrs; 11 sites total): filter to get only LS, TSF more than 10 years, keep only species and cover columns
bird_dat_LS2 <- bird_dat_long %>% 
  filter(sev == "l")

bird_dat_LS2$fire_yr <- as.character(bird_dat_LS2$fire_yr)
bird_dat_LS2$fire_yr <- as.numeric(bird_dat_LS2$fire_yr)
bird_dat_LS2$tsf <- 2018-bird_dat_LS2$fire_yr

bird_dat_LS2 <- bird_dat_LS2 %>% 
  filter(tsf > 15) %>% 
  dplyr::select(species, pa, site_id) %>% 
  arrange(species)
  
bird_dat_LS2 <- reshape(bird_dat_LS2, idvar = "species", timevar = "site_id", direction = "wide")

# add species not detected to get same length vectors
xx.w <- full_join(bird_spp, bird_dat_LS2, by = "species") %>% 
  replace(., is.na(.), "0")

# HS data (less than 11 yrs; 13 sites): filter to get only LS, TSF less than 10 years, keep only species and cover columns
bird_dat_HS1 <- bird_dat_long %>% 
  filter(sev == "h")

bird_dat_HS1$fire_yr <- as.character(bird_dat_HS1$fire_yr)
bird_dat_HS1$fire_yr <- as.numeric(bird_dat_HS1$fire_yr)
bird_dat_HS1$tsf <- 2018-bird_dat_HS1$fire_yr

bird_dat_HS1 <- bird_dat_HS1 %>% 
  filter(tsf < 15) %>% 
  dplyr::select(species, pa, site_id) %>% 
  arrange(species)

bird_dat_HS1 <- reshape(bird_dat_HS1, idvar = "species", timevar = "site_id", direction = "wide")

#bird_dat_LS11 

cm.w <- full_join(bird_spp, bird_dat_HS1, by = "species") %>% 
  replace(., is.na(.), "0")

# HS data (more than 10 yrs; 10 sites): filter to get only LS, TSF more than 10 years, keep only species and cover columns
bird_dat_HS2 <-bird_dat_long %>% 
  filter(sev == "h")

bird_dat_HS2$fire_yr <- as.character(bird_dat_HS2$fire_yr)
bird_dat_HS2$fire_yr <- as.numeric(bird_dat_HS2$fire_yr)
bird_dat_HS2$tsf <- 2018-bird_dat_HS2$fire_yr

bird_dat_HS2 <- bird_dat_HS2 %>% 
  filter(tsf > 15) %>% 
  dplyr::select(species, pa, site_id) %>% 
  arrange(species)

bird_dat_HS2 <- reshape(bird_dat_HS2, idvar = "species", timevar = "site_id", direction = "wide")

#bird_dat_HS22 

dl.w <- full_join(bird_spp, bird_dat_HS2, by = "species") %>% 
  replace(., is.na(.), "0")
```

```{r Bird Convert abundance data to presence-absence, echo = FALSE, include = F}
#Only run the following chunk if you need to convert abundance data to presence-absence.

convert <- function (x) {
  species <- x %>% dplyr::select (species)
  s <- x %>% dplyr::select(-species)
  df <-  data.frame(t(apply (as.matrix(s), 1, function (x) ifelse (x>0, 1, 0))))
  b <- bind_cols (species, df)
  return (b)}

ar.w <- convert (ar.w)
be.w <- convert (be.w)
xx.w <- convert (xx.w)
cm.w <- convert (cm.w)
dl.w <- convert (dl.w)
```
 

```{r Bird Melt data, warning=FALSE, results="hide", echo = FALSE, include = F}
#Convert the data format from wide to long (such that there are only three columns: species, site, value), and order by species. Then delete the site column. 

melt.sort <- function (x){
  m <- melt(x, id.vars="species") # convert to long
  s <- m[order(m$species),] # order by species
  s$variable <- NULL # delete site column
  return(s)}

# l = data are in long format
ar.l <- melt.sort(ar.w) %>% 
  mutate(value = as.numeric(value))
be.l <- melt.sort(be.w) %>% 
  mutate(value = as.numeric(value))
xx.l <- melt.sort(xx.w) %>% 
  mutate(value = as.numeric(value))
cm.l <- melt.sort(cm.w) %>% 
  mutate(value = as.numeric(value))
dl.l <- melt.sort(dl.w) %>% 
  mutate(value = as.numeric(value))

```

```{r Bird Presence-absence, echo= FALSE, include = F}
#### Presence-absence data

#Use this method for presence-absence datasets which don't contain subsamples.

res.pa <- function (z) {(ddply (z, .(species), function(x) { 
  mu <- mean (x$value)
  rn <- (rbinom (n = 1, size = length(x$value), prob = mu))/length(x$value) 
  # rn = random number based on mu
  data.frame(rn=rn)}
  ))
}

# Run 'res.pa' on each growth stage:
ar.r <- rdply (1000, res.pa (ar.l))
be.r <- rdply (1000, res.pa (be.l))
xx.r <- rdply (1000, res.pa (xx.l))
cm.r <- rdply (1000, res.pa (cm.l))
dl.r <- rdply (1000, res.pa (dl.l))

# Combine to a matrix:
zdata <- cbind (ar.r$.n, ar.r$rn, be.r$rn, xx.r$rn, cm.r$rn, dl.r$rn) # ar.r$.n = row labels
```

```{r Combine bird data, echo = FALSE, include = F}
### Convert the matrix to a data frame and save it.
#The same method applies to all data types from this point.
mdata <- apply (zdata, 2, function (x) ifelse (x>0, x, 0.001)) 
# replace zeros for geometric mean calculation
rdata <- data.frame(mdata) # convert to a data frame
names(rdata) <- c("group", "AR", "BE", "XX", "CM", "DL")
rdata$group <- factor(rdata$group) # rdata is ready for the GSO function
```

```{r GSO function bird, echo = FALSE, warning=FALSE}
library(nloptr)
library(plyr)
library(dplyr)
## Handy helper function, constructing a n-length vector c(k, ..., k)
const.vector <- function(n, k) {
  return( sapply(seq_len(n), function (x) return(k)) )
}
## The original objective (assumes columns [AR, BE, CM, DL] appear in that order) 
geomean.fun <- function(x, species) {
  logsum <- sum(log(apply(species, 1, function (row) { sum(x*row) })))
  exp(logsum/(dim(species)[1]))
}
## The reformulated objective (nloptr expects a minimization problem)
geomean.obj <- function (x, species) {
  logsum <- sum(log(apply(species, 1, function (row) { sum(x*row) })))
  return(-logsum)
}
## The gradient of geomean.obj.
## Returns a vector containing the partial derivatives wrt each var.
geomean.grad <- function(x, species) {
  contrib <- apply(species, 1, function(row) { -row/(sum(x*row)) })
  grad <- apply(contrib, 1, sum)  # sum the contribution vectors to get the gradient
  return(grad)
}
## Inequality constraints.
eval.cs <- function(x, species) {
  return( c(
    sum(x) - 1,  # Growth stage proportions don't exceed 1
    0.0 - x[1]   # Recently burnt proportion is at least 0.0 (i.e. AR >= 0.0)
  ))
}
eval.cs.jac <- function(x, species) { # Jacobian (partial derivative matrix) for inequalities
  return( rbind(
    const.vector(length(x), 1),
    c(-1, const.vector(length(x)-1, 0))
    ))
}
## Pass all to nloptr 
gso <- function(species){
  require(nloptr)
  require(dplyr)
  x0 <- const.vector(5, 1/5)  # edit according to the number of growth stages
  run <- nloptr(x0 = x0, species = species, eval_f=geomean.obj, eval_grad_f = geomean.grad,
                lb = const.vector(length(x0), 0),
                ub = const.vector(length(x0), 1),
                eval_g_ineq = eval.cs,
                eval_jac_g_ineq = eval.cs.jac,
                opts = list("algorithm" = "NLOPT_LD_MMA",
                            "maxeval"=1000,
                            "ftol_rel"=1.0e-15,
                            "xtol_rel"=1.0e-8,
                            "print_level"=0
                            ))
  res <- rbind(c(run$solution, geomean.fun(run$solution, species)))
  colnames(res) <- c("AR", "BE", "XX", "CM", "DL", "geom") # edit according to growth stages
  res <- data.frame(res)
  return (res)
}

# Apply to the 1000 resampled communities (defined by group)
outb <- ddply(rdata[2:6], .(rdata$group), gso) # takes a few minutes to run

# Or apply to a single community (for example, the raw data)
#out <- gso (rdata)

#write.table (out, "gso_result.txt", col.names=TRUE, row.names=FALSE)


names(outb)[1]<-"num"
names(outb)[2]<-"UN"
names(outb)[3]<-"LS1"
names(outb)[4]<-"LS2"
names(outb)[5]<-"HS1"
names(outb)[6]<-"HS2"

UN <- mean(outb$UN) *100
LS1 <- mean(outb$LS1) *100
LS2 <- mean(outb$LS2) *100
HS1 <- mean(outb$HS1) *100
HS2 <- mean(outb$HS2) *100

GSO_tbl <- list(UN, LS1, LS2, HS1, HS2) 

names(GSO_tbl)[1]<-"UN"
names(GSO_tbl)[2]<-"LS1"
names(GSO_tbl)[3]<-"LS2"
names(GSO_tbl)[4]<-"HS1"
names(GSO_tbl)[5]<-"HS2"

GSO_tbl <- data.frame(GSO_tbl)

#UN <- sd(out$UN) *100
#LS1 <- sd(out$LS1) *100
#LS2 <- sd(out$LS2) *100
#HS1 <- sd(out$HS1) *100
#HS2 <- sd(out$HS2) *100

#kable(GSO_tbl, caption = "Table 6: GSO results (birds)", digits = 0, col.names = c("UN", "LS1", "LS2", "HS1", "HS2")) %>%
#  kable_styling()

out_long_bird <- gather(outb, severity) %>% 
  filter(severity != "num" & severity != "geom") 

out_long_bird$value <- as.numeric(out_long_bird$value)

out_long_bird <- out_long_bird %>% 
  mutate("prop" = value*100)

# summarize data
data_summary <- function(data, varname, groupnames){
  require(plyr)
  summary_func <- function(x, col){
    c(mean = mean(x[[col]], na.rm=TRUE),
      sd = sd(x[[col]], na.rm=TRUE))
  }
  data_sum<-ddply(data, groupnames, .fun=summary_func,
                  varname)
  data_sum <- rename(data_sum, c("mean" = varname))
 return(data_sum)
}

df3 <- data_summary(out_long_bird, varname="prop", groupnames="severity")

#plot
hist <- ggplot(df3, aes(x = severity, y = prop)) +
  geom_histogram(stat = "identity") +
  labs(title = "Figure 4: Optimal habitat proportions (birds)", y = "% of landscape", x = "habitat") +
  geom_errorbar(aes(ymin=prop-sd, ymax=prop+ sd), width=.2) 
plot(hist)
```


#### GSO: insects
```{r Read in insect data, echo=FALSE, include=F}
## Packages needed throughout
#library (plyr)
#library (dplyr)
#library (reshape2)

#setwd("C:/Users/Documents/Example")

# w = data are in wide format; a, b, c, d keep growth stages in order

#select just the species column
insect_dat_long <- insect_mat %>%
  left_join(site_data, by = "site_id") %>% 
  gather(key = "taxon", value = "number", 3:19)# %>% 
#  dplyr::select(site_id, taxon, number)

#now make a list of all species
#insect_dat <- read.csv("data/insect_data_natives.csv", header=TRUE)

insect_dat_UN <- insect_dat_long %>% 
  filter(sev == "u") %>% 
  dplyr::select(taxon)

insect_spp <- insect_dat_long %>% 
  dplyr::select(taxon)

complete_spp <- union(insect_spp, insect_dat_UN) # but should the dummy species be removed?

#insect_dat_wide <- spread(insect_dat_long, key = "taxon", value = "number")

## Unburnt data: filter to get only UN; keep only species and cover columns
insect_dat_UN <- insect_dat_long %>% 
  filter(sev == "u") %>% 
  dplyr::select(taxon, number, site_id)  %>% 
  arrange(taxon)

#insect_dat_UN <- full_join(insect_dat_UNa, insect_dat_UNb, by = "species")%>% 
#  arrange(species)

# make it wide format
insect_dat_UN <- reshape(insect_dat_UN, idvar = "taxon", timevar = "site_id", direction = "wide")

# add species not detected to get same length vectors
#insect_dat_UN1 
ar.w <- full_join(complete_spp, insect_dat_UN, by = "taxon") %>% 
  replace(., is.na(.), "0")

## LS data (less than 14 yrs; ): filter to get only LS, TSF less than 14 years, keep only species and cover columns
insect_dat_LS1 <- insect_dat_long %>% 
  filter(sev == "l")

insect_dat_LS1$fire_yr <- as.character(insect_dat_LS1$fire_yr)
insect_dat_LS1$fire_yr <- as.numeric(insect_dat_LS1$fire_yr)
insect_dat_LS1$tsf <- 2018-insect_dat_LS1$fire_yr

insect_dat_LS1 <- insect_dat_LS1 %>% 
  filter(tsf < 11) %>% 
  dplyr::select(taxon, number, site_id) %>% 
  arrange(taxon)
  
insect_dat_LS1 <- reshape(insect_dat_LS1, idvar = "taxon", timevar = "site_id", direction = "wide")

# add species not detected to get same length vectors
#insect_dat_LS1 

be.w <- full_join(complete_spp, insect_dat_LS1, by = "taxon") %>% 
  replace(., is.na(.), "0")

## LS data (less than 10 yrs; 11 sites total): filter to get only LS, TSF more than 10 years, keep only species and cover columns
insect_dat_LS2 <- insect_dat_long %>% 
  filter(sev == "l")

insect_dat_LS2$fire_yr <- as.character(insect_dat_LS2$fire_yr)
insect_dat_LS2$fire_yr <- as.numeric(insect_dat_LS2$fire_yr)
insect_dat_LS2$tsf <- 2018-insect_dat_LS2$fire_yr

insect_dat_LS2 <- insect_dat_LS2 %>% 
  filter(tsf > 10) %>% 
  dplyr::select(taxon, number, site_id) %>% 
  arrange(taxon)
  
insect_dat_LS2 <- reshape(insect_dat_LS2, idvar = "taxon", timevar = "site_id", direction = "wide")

# add species not detected to get same length vectors
#insect_dat_LS1 

xx.w <- full_join(complete_spp, insect_dat_LS2, by = "taxon") %>% 
  replace(., is.na(.), "0")

# HS data (less than 14 yrs; ?? sites): filter to get only LS, TSF less than 14 years, keep only species and cover columns
insect_dat_HS1 <- insect_dat_long %>% 
  filter(sev == "h")

insect_dat_HS1$fire_yr <- as.character(insect_dat_HS1$fire_yr)
insect_dat_HS1$fire_yr <- as.numeric(insect_dat_HS1$fire_yr)
insect_dat_HS1$tsf <- 2018-insect_dat_HS1$fire_yr

insect_dat_HS1 <- insect_dat_HS1 %>% 
  filter(tsf < 11) %>% 
  dplyr::select(taxon, number, site_id) %>% 
  arrange(taxon)

insect_dat_HS1 <- reshape(insect_dat_HS1, idvar = "taxon", timevar = "site_id", direction = "wide")

#insect_dat_LS11 

cm.w <- full_join(complete_spp, insect_dat_HS1, by = "taxon") %>% 
  replace(., is.na(.), "0")

# HS data (more than 14 yrs; ?? sites): filter to get only LS, TSF more than 14 years, keep only species and cover columns
insect_dat_HS2 <- insect_dat_long %>% 
  filter(sev == "h")

insect_dat_HS2$fire_yr <- as.character(insect_dat_HS2$fire_yr)
insect_dat_HS2$fire_yr <- as.numeric(insect_dat_HS2$fire_yr)
insect_dat_HS2$tsf <- 2018-insect_dat_HS2$fire_yr

insect_dat_HS2 <- insect_dat_HS2 %>% 
  filter(tsf > 10) %>% 
  dplyr::select(taxon, number, site_id) %>% 
  arrange(taxon)

insect_dat_HS2 <- reshape(insect_dat_HS2, idvar = "taxon", timevar = "site_id", direction = "wide")

#insect_dat_HS22 

dl.w <- full_join(complete_spp, insect_dat_HS2, by = "taxon") %>% 
  replace(., is.na(.), "0")
```

```{r Melt insect data, warning=FALSE, results="hide", echo =F, include = F}
#Convert the data format from wide to long (such that there are only three columns: taxon, site_id, number), and order by taxon Then delete the site column. 

melt.sort <- function (x){
  m <- melt(x, id.vars="taxon") # convert to long
  s <- m[order(m$taxon),] # order by species
  s$variable <- NULL # delete site column
  return(s)}

# l = data are in long format
ar.l <- melt.sort(ar.w) %>% 
  mutate(value = as.numeric(value))
be.l <- melt.sort(be.w) %>% 
  mutate(value = as.numeric(value))
xx.l <- melt.sort(xx.w) %>% 
  mutate(value = as.numeric(value))
cm.l <- melt.sort(cm.w) %>% 
  mutate(value = as.numeric(value))
dl.l <- melt.sort(dl.w) %>% 
  mutate(value = as.numeric(value))

#Now data are ready for resampling.  Use a different method for abundance data (resampling with replacement) and presence-absence data (random number generation from the binomial distribution).  Both methods produce a matrix named `zdata`.  

#An extra step is required for data containing subsamples -- see below.

```


```{r insect Abundance, echo=FALSE, include = F}
#### Abundance data
#Use this method for abundance datasets which don't contain subsamples.

res.ab <- function (z){(ddply (z, .(taxon), function(x) { 
  rs1 <- sample(x$value, replace = TRUE) # resample raw data
  bm <- mean(rs1) # big mean is the mean of the resampled means  
  data.frame(bm=bm)}
  ))
}

# Run 'res.ab' on each growth stage:
ar.r <- rdply (1000, res.ab (ar.l))
be.r <- rdply (1000, res.ab (be.l))
xx.r <- rdply (1000, res.ab (xx.l))
cm.r <- rdply (1000, res.ab (cm.l))
dl.r <- rdply (1000, res.ab (dl.l))

# Combine to a matrix:
zdata <- cbind (ar.r$.n, ar.r$bm, be.r$bm, xx.r$bm, cm.r$bm, dl.r$bm) # ar.r$.n = row labels
```

```{r Combine insect data, echo = FALSE,include =F}
### Convert the matrix to a data frame and save it.
#The same method applies to all data types from this point.

mdata <- apply (zdata, 2, function (x) ifelse (x>0, x, 0.001)) 
# replace zeros for geometric mean calculation
rdata <- data.frame(mdata) # convert to a data frame
names(rdata) <- c("group", "AR", "BE", "XX", "CM", "DL")
rdata$group <- factor(rdata$group) # rdata is ready for the GSO function
```


```{r GSO function for insects, echo = FALSE, warning=FALSE, message = F}
## 2 Run the GSO function

#The nloptr function maximises the geometric mean by changing the growth stage proportions given bounds constraints (parameters (GS proportions) must be between 0 and 1), and an inequality constraint -- the parameters must sum to 1. 

## Handy helper function, constructing a n-length vector c(k, ..., k)
const.vector <- function(n, k) {
  return( sapply(seq_len(n), function (x) return(k)) )
}
## The original objective (assumes columns [AR, BE, CM, DL] appear in that order) 
geomean.fun <- function(x, taxon) {
  logsum <- sum(log(apply(taxon, 1, function (row) { sum(x*row) })))
  exp(logsum/(dim(taxon)[1]))
}
## The reformulated objective (nloptr expects a minimization problem)
geomean.obj <- function (x, taxon) {
  logsum <- sum(log(apply(taxon, 1, function (row) { sum(x*row) })))
  return(-logsum)
}
## The gradient of geomean.obj.
## Returns a vector containing the partial derivatives wrt each var.
geomean.grad <- function(x, taxon) {
  contrib <- apply(taxon, 1, function(row) { -row/(sum(x*row)) })
  grad <- apply(contrib, 1, sum)  # sum the contribution vectors to get the gradient
  return(grad)
}
## Inequality constraints.
eval.cs <- function(x, taxon) {
  return( c(
    sum(x) - 1,  # Growth stage proportions don't exceed 1
    0.0 - x[1]   # Recently burnt proportion is at least 0.0 (i.e. AR >= 0.0)
  ))
}
eval.cs.jac <- function(x, taxon) { # Jacobian (partial derivative matrix) for inequalities
  return( rbind(
    const.vector(length(x), 1),
    c(-1, const.vector(length(x)-1, 0))
    ))
}
## Pass all to nloptr 
gso <- function(taxon){
  require(nloptr)
  require(dplyr)
  x0 <- const.vector(5, 1/5)  # edit according to the number of growth stages
  run <- nloptr(x0 = x0, taxon = taxon, eval_f=geomean.obj, eval_grad_f = geomean.grad,
                lb = const.vector(length(x0), 0),
                ub = const.vector(length(x0), 1),
                eval_g_ineq = eval.cs,
                eval_jac_g_ineq = eval.cs.jac,
                opts = list("algorithm" = "NLOPT_LD_MMA",
                            "maxeval"=1000,
                            "ftol_rel"=1.0e-15,
                            "xtol_rel"=1.0e-8,
                            "print_level"=0
                            ))
  res <- rbind(c(run$solution, geomean.fun(run$solution, taxon)))
  colnames(res) <- c("AR", "BE", "XX", "CM", "DL", "geom") # edit according to growth stages
  res <- data.frame(res)
  return (res)
}

# Apply to the 1000 resampled communities (defined by group)
out <- ddply(rdata[2:6], .(rdata$group), gso) # takes a few minutes to run

names(out)[1]<-"num"
names(out)[2]<-"UN"
names(out)[3]<-"LS1"
names(out)[4]<-"LS2"
names(out)[5]<-"HS1"
names(out)[6]<-"HS2"

UN <- mean(out$UN) *100
LS1 <- mean(out$LS1) *100
LS2 <- mean(out$LS2) *100
HS1 <- mean(out$HS1) *100
HS2 <- mean(out$HS2) *100

GSO_tbl <- list(UN, LS1, LS2, HS1, HS2) 

names(GSO_tbl)[1]<-"UN"
names(GSO_tbl)[2]<-"LS1"
names(GSO_tbl)[3]<-"LS2"
names(GSO_tbl)[4]<-"HS1"
names(GSO_tbl)[5]<-"HS2"

GSO_tbl <- data.frame(GSO_tbl)

#kable(GSO_tbl, caption = "Table 5: GSO results (insects)", digits = 0, col.names = c("UN", 
#                                                                           "LS1", 
#                                                                           "LS2", 
#                                                                           "HS1", 
#                                                                           "HS2")) %>%
#  kable_styling()

out_long <- gather(out, severity) %>% 
  filter(severity != "num" & severity != "geom") 

out_long$value <- as.numeric(out_long$value)

out_long <- out_long %>% 
  mutate("prop" = value*100)

# summarize data
data_summary <- function(data, varname, groupnames){
  require(plyr)
  summary_func <- function(x, col){
    c(mean = mean(x[[col]], na.rm=TRUE),
      sd = sd(x[[col]], na.rm=TRUE))
  }
  data_sum<-ddply(data, groupnames, .fun=summary_func,
                  varname)
  data_sum <- rename(data_sum, c("mean" = varname))
 return(data_sum)
}

df2 <- data_summary(out_long, varname="prop", groupnames="severity")

#plot
hist <- ggplot(df2, aes(x = severity, y = prop)) +
  geom_histogram(stat = "identity") +
  labs(title = "Figure 3: Optimal habitat proportions (insects)", y = "% of landscape", x = "habitat") +
  geom_errorbar(aes(ymin=prop-sd, ymax=prop+ sd), width=.2) 
hist


# Or apply to a single community (for example, the raw data)
#out <- gso (rdata)

#write.table (out, "gso_result.txt", col.names=TRUE, row.names=FALSE)
```

# Discussion and next steps  

My data show that high severity areas are important for biodiversity compared to fire-suppressed stands, even those that have burnt once at low severity. HS burns provide habitat to numerous species not found in fire-suppressed stands, and the GSO reflects this by suggesting a vast proportion of the landscape should be burnt at high severity to achieve the highest biodiversity levels.  

But although this result can be interpreted to mean that high severity fire is a normal and positive element in the fire regime of the Klamath Mountains and that a hands-off fire management approach is best, an alternative explanation is worth considering. Instead of highlighting the value of high severity burns, my data might be showing the **dramatic cost to biodiversity of fire suppression**. Historically, numerous species that occur today in high severity areas would likely have used open-canopied forests maintained by a more frequent fire regime. In such habitats, these species would have coexisted with those that require mature stands and partial canopy cover. Conversion to closed-canopy forests may have made vast areas unsuitable for open canopy species which now find themselves pushed into the only remaining habitat they can use, high severity burns.  

The resulting dichotomy between HS burns and closed-canopy stands may be a novel condition on the landscape. While calls for forest restoration are often concerned with reducing the risk of catastrophic fire, my data suggest that **the real catastrophe from a biodiversity perspective is the widespread canopy closure caused by fire suppression**. Therefore, rather than celebrating HS fire and vindicating advocates of hands-off approaches to fire management, my data might be pointing to the urgent need to restore closed canopy, fire-suppressed forests to their historical condition.  

Needless to say, my sampling methods had their limitations as they only focussed on a subset of species. I did not survey many groups of organisms that might be tied to high canopy cover and mature stands, including bryophytes, molluscs, and amphibians. Additionally, a focus on species diversity in high severity burns ignores negative impacts of such fires, such as sedimentation, loss of legacy trees, and numerous cultural values. These matters would be discussed at greater length in the final product.  

Additional research is needed to answer some questions that are prompted by these results:     
      
+   A single low severity fire seems to have little impact on biodiversity, but does frequent LS fire have a significant impact? Does frequent LS fire lead to lower canopy cover over time and does this enable species affiliated with HS burns to coexist within the same stand with species affiliated with closed-canopy stands?     
+   What was the historical understory like and how does it compare to what we see today in either high or low severity burns?  

The first question can be answered by studying areas that have burnt multiple times, probably using prescribed fire. Though few such areas exist, using sites further afield (eg. the Rogue-Siskiyou National Forest) might be sufficient for such a study. Since my data suggest little benefit to biodiversity from a single burn, it would be interesting to see how management using multiple prescribed burns can achieve biodiversity enhancement goals.  

The second question can be answered using phytolith analysis, which has not been done in the Klamath Mountains to my knowledge. However, I have no experience with this method, and it might be unwise to start down a new path (if I want to finish my degree in a timely manner), but the results might be worth the effort. I will investigate collaborating with experienced phytolith researchers at UC Berkeley.  
  
    
      
        
        
        
          

<a href="#header">back to top</a>  

****************************************
