---
title: "Methods"
author: "Chris Adlam"
header-includes:
 - \usepackage{wrapfig}
 - \usepackage{lipsum}
output:
 pdf_document:
  keep_tex: true
 html_document: default
 word_document: default
# extra_dependencies: flafter
bibliography: bibliography3.bib
editor_options:
 chunk_output_type: console
---

```{r, include = F}
# explanations at https://stackoverflow.com/questions/54100806/r-markdown-how-do-i-make-text-float-around-figures/54112408#54112408
defOut <- knitr::knit_hooks$get("plot") # save the default plot hook 
knitr::knit_hooks$set(plot = function(x, options) { # set new plot hook ...
 x <- defOut(x, options) # first apply the default hook
 if(!is.null(options$wrapfigure)) { # then, if option wrapfigure is given ...
  # create the new opening string for the wrapfigure environment ...
  wf <- sprintf("\\begin{wrapfigure}{%s}{%g\\textwidth}", options$wrapfigure[[1]], options$wrapfigure[[2]])
  x <- gsub("\\begin{figure}", wf, x, fixed = T) # and replace the default one with it.
  x <- gsub("{figure}", "{wrapfigure}", x, fixed = T) # also replace the environment ending
 }
 return(x)
})
```


# Methods

## Study area
Located in the Klamath-Siskiyou mountains of northern California, the WKRP is a 1.2 million-acre collaborative landscape restoration effort between the Karuk Tribe, US Forest Service and local NGOs [@harlingWesternKlamathRestoration2014] (Figure 1). The area is dominated by mixed evergreen forest [@steelFireFrequencyseverityRelationship2015], which is characterized by the prevalence of Douglas-fir (*Pseudotsuga menziesii*) and diverse hardwoods such as tanoak (*Notholithocarpus densiflorus*), black oak (*Quercus kelloggii*), canyon live oak (*Quercus chrysolepis*), and madrone (*Arbutus menziesii*) [@barbourNorthAmericanTerrestrial2000; @whittakerVEGETATIONHISTORYPACIFIC1961]. The landscape of the study area was historically managed by the Karuk tribe using frequent (~12.5 year interval), low-intensity prescribed burning [@taylorFireHistoryLandscape1998; @taylorSpatialPatternsControls2003a; @crawfordLateHoloceneFire2015]; lakeHistoricalCulturalFires2013]. As a result, the historic fire regime was one of more frequent and generally lower severity fires than would have been expected under the influence of climate alone [@taylorFireHistoryLandscape1998; @carloniEcologicalLegacyIndian2005; @skinnerHumanClimaticInfluences2009a; @taylorSocioecologicalTransitionsTrigger2016]. One of the main goals of the WKRP is to restore the fire regimes within the Karuk tribe’s aboriginal territory in a way that promotes the resilience of these ecosystems to high-severity fire and climate change, and increases biodiversity and the provision of diverse ecosystem services [@harlingWesternKlamathRestoration2014].

## Data collection
I conducted surveys at 110 sites varying in time-since-fire and burn severity (48 sites surveyed in 2018 and 62 in 2019). Sampling sites were in lower elevation (mean 759m, range 241-1500m) mixed evergreen forest, with some in transition areas into dry or moist mixed conifer (classification based on presettlement fire regime [@vandewaterSummaryFireFrequency2011], accessed through the FRID dataset [@saffordCaliforniaFireReturn2011]). The median historic fire return interval in these forests is 13 years [@vandewaterSummaryFireFrequency2011], with some variation based on aspect and other topographic and climatlic factors [@taylorFireHistoryLandscape1998; @skinnerKlamathMountainsBioregion2006].


```{r echo = F, warning = F, message = F, fig.width=7, fig.height = 5, out.width = ".5\\textwidth", fig.cap = "Map of WKRP project area and sampling sites", fig.align="right", wrapfigure = list("R", .5)}
include_graphics2 <- function(path, alt_path = NULL, handler = function(path) knitr::asis_output(paste('View', tools::file_ext(path), 'at', path)), ...) {
 if (knitr::is_latex_output()) {
  return(include_graphics_latex(path, alt_path, handler, ...))
 } else {
  return(knitr::include_graphics(path, ...))
 }
}

#' Wrapper of \code{knitr::include_graphics} for PDF Output
#'
#' Deals with URL and GIFs. If an url is passed to
#' \code{path} of \code{\link[knitr]{include_graphics}},
#' the figure is automatically downloaded and included
#' using local relative path. If a figure with \code{.gif}
#' extension is included, a piece of text, rather than the
#' figure, is inserted.
#'
#' @importFrom tools file_ext file_path_sans_ext
#' @importFrom utils download.file
#' @importFrom knitr current_input
#' @keywords internal
include_graphics_latex <- function(path, alt_path = NULL, handler = function(path) knitr::asis_output(paste('View', tools::file_ext(path), 'at', path)), ...) {
 # URL
 if (grepl('^https?://', path)) {
   ifelse(use_alt_path(path, alt_path),
      path <- alt_path,
      return(handler(path)))

  ## Download Figure
  dir_path <- paste0('downloadFigs4latex_',
            file_path_sans_ext(current_input()))
  if (!dir.exists(dir_path)) dir.create(dir_path)
  file_path <- paste0(dir_path, '/',
            knitr::opts_current$get()$label, '.',
            file_ext(path))
  download.file(path, destfile = file_path)
  path <- file_path
 }

 # Local files
 else {
   ifelse(use_alt_path(path, alt_path),
      path <- alt_path,
      return(handler(path)))
 }

 # Insert Figure
 return(knitr::include_graphics(path, ...))
}


use_alt_path <- function(path, alt_path) {
 # Invalid img ext & no alt provided: Don't include in File
 if (inval_latex_img(path) && is.null(alt_path)) return(FALSE)
 # Invalid img ext with alt provided: insert alt-figure
 if (inval_latex_img(path) && !is.null(alt_path)) {
   stopifnot(!inval_latex_img(alt_path))
   return(TRUE)
  }
}

inval_latex_img <- function(path) {
 invalid_ext <- c('svg', 'SVG', 'GIF', 'gif')
 return(tools::file_ext(path) %in% invalid_ext)
}

include_graphics2("/Users/christopheradlam/Desktop/Davis/R/GitHub Repos/Fire_mosaics/output/WKRP_map.png")
```

Sampling sites were placed in low-severity ("LS") and high-severity burns ("HS"), defined respectively as <25% canopy mortality and >75% canopy mortality [@millerQuantitativeEvidenceIncreasing2009], in long-unburnt stands ("UN", >80 years since last fire), in sites affected by multiple burns ("MULT"), and sites that were thinned and burnt ("Rx"). At all burnt sites, time-since-fire varied from 2 to 32 years. Multiple burns and thinned and burnt sites were placed in areas with intermediate canopy cover (30-75% cover). While this eliminates high-severity reburns and multiple burns or restoration treatments that retain high levels of canopy cover, this study only considered "treatments" (active or passive) that reduce canopy cover significantly. Such canopy reduction corresponds to the level recommended for understory restoration [@abellaEffectsTreeCutting2015] and for mitigating the risk of crown fires [@moghaddasFuelTreatmentEffects2010]. As a result of this approach, the habitats studied follow a gradient of canopy cover (Figure 2). Burn severity was evaluated a priori based on the relative difference normalized burn ratio [@millerQuantifyingBurnSeverity2007; @millerCalibrationValidationRelative2009] and verified in situ by visually estimating the magnitude of canopy cover reduction. Since the habitats differ mostly by the severity of fire's impact, I will henceforth refer to different "severity categories" for simplicity, even though this is an imperfect description of multiple burn and thinned and burnt sites.

Several tests were used to ensure that the differences observed between severity categories were not simply the product of a climatic or topographic pattern. To test for climate differences, I compared the climatic water deficit (the difference between potential and actual evapotranspiration). Using the Basin Characterization Model 30-year average for 1981-2010 [@flintFinescaleHydrologicModeling2013], I found no difference between the severity categories (ANOVA, *F*~(4,105)~ = 1.41, *p* = 0.237). Then, I used measurements of slope, aspect and latitude to calculate the heat load at each site. Heat load indicates the magnitude of insolation and temperature and is equivalent to solar radiation but for two differences: it is symmetrical along the northeast-southwest axis rather than the north-south axis (since a slope with afternoon sun will have higher maximal temperatures than a similar slope with morning sun), and it can be calculated using a simple equation instead of requiring simulation modeling or manual interpolation using solar radiation tables [@mccuneEquationsPotentialAnnual2002]. I found no detectable difference among categories except that heat load was slightly lower for high-severity sites (although this might seem counterintuitive, it is explained by the fact that high-severity sites were steeper on average, and heat load decreases with slope gradient on north-facing aspects). However, due to the limited availability of thinned and burnt areas, I had to make some concessions in study design. While there were numerous sites available for the other severity categories, the only prescribed burn sites that met this study's criteria for minimum size, degree of canopy reduction, and absence of other disturbance factors were located in two project areas. These were higher in elevation on average than the other severity categories (mean = 1101m, 95% CL = 965-1237m) and more clustered (average distance among sites <10km compared to 28-36km for the other severity categories). Spatial autocorrelation of fire effects among plots within a burn unit is uncommon [@vanmantgemNegligibleInfluenceSpatial2009], but ecological communities could be expected to be more similar as a result of these sampling limitations. While these factors limit the generalizability of the patterns observed at these sites, this is a common challenge for studies of prescribed fires. Therefore I still include the thin+burn sites in the study in the hope that my findings can be used in designing more robust future investigations.

At each site, I recorded aspect, slope, elevation, and canopy cover (trees over 5m in height, by visual estimation). For all sites, I surveyed plants, epiphytic lichens, and birds. Presence of all plants and lichen species was determined in an 11.3m-radius plot (400 m^2^), a size commonly used for Common Stand Exams [@usfsCommonStandExam2008]. Plants were identified using the Jepson manual [@baldwinJepsonManualVascular2012] and lichens were identified using Mccune and Geiser's Macrolichens of the Pacific Northwest [@mccuneMacrolichensPacificNorthwest2009]. Only native plant species are included in the analysis, although exotic species were generally few. Species of some plant genera were pooled if they shared similar ecological requirements (eg. *Arctostaphylos*, *Cryptantha*, *Pyrola*). Species of the lichen genera *Usnea* and *Bryoria* were also pooled because of the difficulty in differentiating them. I also conducted two 10-minute bird point counts at each site using standard methods [@ralphHandbookFieldMethods1993]. These surveys took place on separate, non-consecutive days with no precipitation and little to no wind, between sunrise and 10:00am during breeding season (June-early July). All bird visual and aural detections within a 100m radius were recorded. The two hummingbird species (Anna's and Rufous Hummingbird) were pooled, as were Black-throated Gray Warblers and Hermit Warblers (*Setophaga* sp.), because it was not always possible to differentiate these species by ear. Lastly, for high-severity, low-severity and unburnt sites, I set up a custom-built flight-intercept trap for flying insects that was left out for two days [@russoCompositeInsectTrap2011]. Insects were identified to the order or suborder level.

## Analysis
### Species composition
I conducted a non-parametric multivariate permutational analysis of variance (PERMANOVA) to test the significance of dissimilarity between the different severity categories [@mcardleFittingMultivariateModels2001] using the adonis2() function from the vegan package v.2.5.6 in R [@oksanenVeganPackage2007], with 999 permutations. PERMANOVA calculates a pseudo-F value similar to the F-statistic in ANOVA models, and obtains p-values based on permutation procedures [@andersonPermutationalMultivariateAnalysis2017]. Then I conducted PERMANOVA pairwise comparisons between each severity category. To minimize the risk of making a type-1 error from carrying out multiple pairwise tests on a single data set, I adjusted the p-values using the Bonferroni correction method [@riceAnalyzingTablesStatistical1989].

To visualize the difference in species composition in each severity category, I used non-metric multidimensional scaling (NMDS) based on Bray-Curtis dissimilarity matrices using the metaMDS() function of the vegan package in R [@oksanenVeganPackage2007].

### Diversity patterns
Next, I analyzed patterns of alpha, beta and gamma diversity for plants, birds, and lichens. Alpha diversity is equivalent to species richness per plot. Because assumptions of normality were often unmet with my data, I performed permutational ANOVAs using the aovp() function of the lmPerm package [@wheelerLmPermPermutationTests2016]. Tukey HSD post-hoc tests were used to evaluate the significance of the difference in richness between severity categories.

To further evaluate the influence of canopy cover on species richness, I used linear regressions with plant and bird richness as the response variable and percent tree cover as the predictor for the lower canopy cover habitat (high-severity burns), intermediate canopy habitats (multiple and thinned and burnt stands), and higher canopy cover habitats (low-severity and unburnt). Data were square root transformed to acheive normality (verified using Shapiro-Wilkes tests). This tiered approach was chosen to isolate the effect of treatment and of canopy cover, since the treatments themselves are the principal driver of canopy cover differences.

I calculated beta diversity, or the amount of variation in species composition and richness between sites within each severity category, using the R package vegetarian v1.2 (Charney and Record, 2009). I used the 1-Jaccard index [@jostPartitioningDiversityIndependent2007] to represent the extent to which species are shared in each pair of samples. I compared results between treatments using standard errors from bootstrapping given by the package vegetarian.

Because the number of samples in each severity category was unequal, simply comparing the total number of species found in each severity category was not an adequate way to evaluate gamma diversity, the size of the species pool for each habitat. Instead, I generated sample-based rarefaction curves [@colwellModelsEstimatorsLinking2012] using package iNEXT v. 2.0.20 [@chaoRarefactionExtrapolationHill2014; @hsiehINEXTPackageRarefaction2016]. This method uses Hill numbers, or effective species numbers based on species richness. Estimates were interpolated from site-based incidence data to account for unequal sample sizes [@colwellModelsEstimatorsLinking2012] and then extrapolated to twice the size of the smallest sample [@chaoRarefactionExtrapolationHill2014]. I compared results between treatments using 95% confidence intervals from bootstrapping.

### Indicator species analysis
Insects were not sampled in the multiple burns and thinned and burnt sites, but I include them nevertheless to compare species communities in the unburnt, low-severity and high-severity burns. Comparisons of alpha, beta and gamma diversity were not as meaningful for insects, because of the small number of taxa (at the order or suborder level). Instead, I used an indicator species analysis using the function multipatt() in the package indicspecies v1.7.8 [@decaceresAssociationsSpeciesGroups2009] to determine which taxa showed a preference for high-severity or low-severity and unburnt stands (the latter were pooled because the PERMANOVA suggested their community composition did not differ). I used the same procedure to identify indicator species for birds and plants in the high-severity and low-severity/unburnt categories.

### Early and late seral species in actively and passively managed stands
Lastly, I wanted to determine if the multiple burns and thinned and burnt stands tended to contain species associated with high-severity burns and/or those associated with low-severity and unburnt stands. Four hypotheses were envisioned: 1) Species preferring high-severity burns also occur frequently in the actively or passively managed stands; 2) Species preferring unburnt stands and low-severity burns also occur frequently in the actively or passively managed stands; 3) Both of these species cohorts occur frequently in the actively or passively managed stands (presumably because they are intermediate in environmental characteristics such as canopy cover); 4) Neither cohort is found in the actively or passively managed stands (presumably because the intermediate conditions are inhospitable for both and instead favor a completely different species assemblage). In the first two cases, active or passive management would generate habitat conditions that are ecologically redundant with high-severity burns or denser mature stands. The third hypothesis would mean that reducing canopy cover for fire management would be most compatible with biodiversity objectives, while the fourth would be the worst case scenario. To determine which of these four hypotheses was correct, I identified species that prefer high-severity burns and species that prefer unburnt stands and low-severity burns by using a simple criterion: if a species was found at least twice as frequently in one habitat than the other, I considered that it exhibited a preference for that habitat. Only species that occurred in more than five sites (18% of the total number of sites) were included. I compared the list of early seral species (found twice as frequently in the high-severity burns) and late seral species (found twice as frequently in the low-severity burns and unburnt stands) with findings from geographically-related studies to ensure that my determination was accurate. For both groups of species, I then determined if each species was also found at least twice as frequently in the multiple burns and/or thinned and burnt stands. To determine which of the four hypotheses above was supported by the data, I evaluated the frequency with which species preferring high-severity burns or low-severity burns and unburnt stands also favored the multiple burns and/or thinned and burnt stands. In addition, for these species, Karuk cultural uses from the published literature were included to illustrate the interconnectedness of biological and cultural diversity [@longUsingCulturallySignificant2020; @schenckKarokEthnobotany1952a].